---
title: "IRW explore"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    
---

# next steps: 
plan: try other models 
simulation data: simulation data from DINA 
use frac20 Q matrix, simulate new data 
simulate data from profile 
under situations 
skill heirachy 

1. focus on a set of simulations 
- that will favor CDM model 
1 ability: 1 easy skill and 1 hard skill 
2 dimensional thetas (each has 1 easy and 1 hard): 4 skills 
- 2 thetas correlated <-- 1
- 2 thetas cor <-- 0

2. 

# Data Preprocessing
## Load Packages
Let's first load the packages that we need for this chapter. 

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
library(dplyr)
library(mirt)
library('ggpubr')
library(Metrics)
library(hash)
library(ggplot2)
library(GGally)
library(gridExtra)
library(rstudioapi)
library(TAM)
library(mgcv)
library(GDINA)
library(MASS)
```

```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")

# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

# Workshop Reference
```{r}
data1 <- read.table(file = "data/data1.dat", header = TRUE)

# Q-matrix
Q1 <- read.table(file = "data/Q1.txt")

# Fit the data using G-DINA model
fit1 <- GDINA(dat = data1, Q = Q1, model = "GDINA")

#summary(fit1)

# Level 1: percent of students mastered attribute 

#coef(fit1)

coef(object = fit1, what = "delta")

# this is so important!!!!!!
#head(personparm(fit1, what = "MAP"))
head(personparm(fit1, what = "mp"))

#plot(fit1, what = "mp", person = c(1,5))

## default is EAP

#extract(fit1, what = "posterior.prob")

#Qval(fit1, method = "wald")

#modelfit(fit1)
```

# Load IRW
```{r}
#reliability 
dataset <- redivis::user("datapages")$dataset("item_response_warehouse",version='v7.0')

dataset_tables <- dataset$list_tables()
```

## Get CDM data
```{r}
##
print(length(dataset_tables))
names<-sapply(dataset_tables,function(x) x$name)
ii<-grep("metadata",names)
names(dataset_tables)<-names
if (length(ii)>0) dataset_tables<-dataset_tables[-ii]   
##
f<-function(table) table$list_variables()
nms<-lapply(dataset_tables,f)

##
f<-function(x) {
    nm<-sapply(x,function(x) x$name)
    any(grepl("^Qmat",nm))
}
test<-sapply(nms,f)
q_data<-dataset_tables[test]
print(length(q_data))
```

# helper functions 
## IMV
```{r}
imv.binary<-function(y, #outcomes
                     p1,#baseline
                     p2 #enhanced
                     ) {
    ##
    ll<-function(x,p) {
        z<-log(p)*x+log(1-p)*(1-x)
        z<-sum(z)/length(x)
        exp(z)
    }    
    loglik1<-ll(y,p1)
    loglik2<-ll(y,p2)
    getcoins<-function(a) {
        f<-function(p,a) abs(p*log(p)+(1-p)*log(1-p)-log(a))
        nlminb(.5,f,lower=0.001,upper=.999,a=a)$par
    }
    c1<-getcoins(loglik1)
    c2<-getcoins(loglik2)
    ew<-function(p1,p0) (p1-p0)/p0
    imv<-ew(c2,c1)
    imv
}
```

## IRT
```{r}
prob.irt <-function(a, b,theta) {
  1 / (1 + exp(-a * (theta - b)))
}
```

## Convert to Q matrix and data 
```{r}
func.convert.Q.data.structure <- function(df){
  Q <- df %>% 
    dplyr :: select(-c(id, resp)) %>% 
    unique() %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
    arrange(item_num) %>% 
     dplyr ::select(-c(item, item_num))
  
  dat <- df %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
     dplyr ::select(item_num, id, resp) %>% 
    arrange(item_num) %>% 
    mutate(item = paste0("V",item_num)) %>% 
     dplyr ::select(-item_num) %>% 
    pivot_wider(names_from = item, values_from = resp) %>% 
    arrange(id)
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- dat
  
  return (tables_list)
}

```

```{r}
func.convert.Q.data.structure.with.held.out <- function(df, heldout_rate){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
     dplyr ::arrange(item_num) %>% 
     dplyr ::select(-c(item, item_num))
  
  df.held.out <- df %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
     dplyr ::select(item_num, id, resp) %>% 
    dplyr :: arrange(item_num) %>% 
    mutate(item = paste0("V",item_num)) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = heldout_rate)) %>% 
     dplyr ::arrange(id)
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
     dplyr ::select(-c(item_num, held_out)) %>% 
    pivot_wider(names_from = item, values_from = resp) %>% 
     dplyr ::arrange(id) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- df.dat %>%  dplyr ::select(-id)
  tables_list[['heldout']] <- df.held.out
  tables_list[['id']] <- df.dat$id
  
  return (tables_list)
}
```

```{r}
func.convert.Q.data.structure.with.held.out.limited.skills <- function(df, heldout_rate, limited_skills){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
     dplyr ::arrange(item_num) %>% 
     dplyr ::select(-c(item, item_num))
  
  df.held.out <- df %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
     dplyr ::select(item_num, id, resp) %>% 
    dplyr :: arrange(item_num) %>% 
    mutate(item = paste0("V",item_num)) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = heldout_rate)) %>% 
     dplyr ::arrange(id)
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
    arrange(item_num) %>% 
     dplyr ::select(-c(item_num, held_out)) %>%
    pivot_wider(names_from = item, values_from = resp) %>% 
     dplyr ::arrange(id) 
  
  filtered_df <- df.dat[rowSums(is.na(df.dat)) <= limited_skills, ]
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- filtered_df %>%  dplyr ::select(-id)
  tables_list[['heldout']] <- df.held.out
  tables_list[['id']] <- filtered_df$id
  
  return (tables_list)
}
```

```{r}
func.get.table <- function(num, coef, q_matrix_long) {
  
  # Initialize an empty list to store the data frames
  tables_list <- list()
  
  for (i in seq(1, num, 1)) {
    
    # Filter the required attributes for the current item from q_matrix_long
    required_attribute <- q_matrix_long %>%
      filter(Value == 1, item == i) %>%
      arrange(Attribute)
    
    # Get the attribute names
    attributes <- required_attribute$Attribute
    
    # Create a dynamic name for the current item (e.g., "Item 1", "Item 2", etc.)
    name <- paste0("Item", " ", i)
    
    # Extract the coefficients for the current item and convert to data frame
    df.coef <- as.data.frame(coef[[name]])
    
    # Get the rownames of the coefficient data frame (the patterns)
    list.combination <- rownames(df.coef)
    
    # Function to convert pattern strings (like "P(000)", "P(100)", etc.)
    convert_pattern <- function(pattern) {
      # Remove "P(" and ")" to isolate the binary part
      binary_string <- gsub("[P()]", "", pattern)
      
      # Split the binary string into individual characters
      binary_vector <- as.numeric(unlist(strsplit(binary_string, "")))
      
      # Return the binary vector
      return(binary_vector)
    }
    
    # Extract probabilities and ensure they are numeric
    possibilities <- as.numeric(df.coef[[1]])
    
    # Convert the list.combination patterns to a matrix form
    attribute_matrix <- do.call(rbind, lapply(list.combination, convert_pattern))
    
    # Convert attribute_matrix to a data frame and ensure numeric
    df <- as.data.frame(attribute_matrix, stringsAsFactors = FALSE)
    df[] <- lapply(df, as.numeric)  # Coerce all columns to numeric
    
    # Add the probabilities as a separate numeric column
    df$Probability <- possibilities
    
    # Assign proper column names
    colnames(df) <- c(attributes, "Probability")
    colnames(df) <- sub("^Qmatrix__", "A", colnames(df))
    
    # Store the resulting data frame in the list
    tables_list[[name]] <- df
    
  }
  
  # Return the list of data frames
  return(tables_list)
}
```


```{r}
func.predict.resp <- function(item_prob, persons){
  persons.prediction <- persons %>% 
    left_join(item_prob) %>% 
    arrange(id)
  return(persons.prediction$Probability)
}

# want to return a dataframe with column i with "name: Vi"and values shall be the list l
func.predict.resp.all.items <- function(df, item, persons, table.list){

  # Initialize an empty list to store predictions
  response_list <- list()
  
  n <- length(df$dat)
  for (i in seq(1, n, 1)) {
    name <- paste0("Item", " ", i)
    
    item_prob <- table.list[[name]]
    
    l <- func.predict.resp(item_prob, persons)
    
    col_name <- paste0("V", i)
    response_list[[col_name]] <- l
  }
  response_df <- as.data.frame(response_list, stringsAsFactors = FALSE)
  
  # Optionally, add identifiers from 'persons' if needed
  # For example, if 'persons' has an ID column:
  # response_df$PersonID <- persons$ID
  
  return(response_df)
}
```


# Analysis (All data)

```{r}
df <- q_data$frac20$to_data_frame()
```

```{r}
df.cdm.structure.frac20 <- func.convert.Q.data.structure.with.held.out(df, 0)
```

fit DINA model

```{r}
mod.dina <- GDINA(df.cdm.structure.frac20$dat,df.cdm.structure.frac20$Q, model = "DINA")

coef <- coef(mod.dina, "itemprob")
```

```{r}
frac20.q <- df.cdm.structure.frac20$Q %>% 
  mutate(item = seq(1, 20, 1))

q_matrix_long <-  frac20.q %>%
  pivot_longer(cols = starts_with("Qmatrix"),  # Select columns that represent attributes
               names_to = "Attribute",  # Name for the new 'attribute' column
               values_to = "Value") 
```

get person estimate

```{r}
persons <- as.data.frame(personparm(mod.dina))


persons.with.id <- persons %>%  
  mutate(id = df.cdm.structure.frac20$id)

#colnames(persons) <- sub("^A", "V", colnames(persons))

name <- paste0("Item", " ", 1)
df.coef <- as.data.frame(coef[[name]])

table.list <- func.get.table(20, coef, q_matrix_long)

n <- length(df.cdm.structure.frac20$dat)
name <- paste0("Item", " ", 1)
item_prob <- table.list[[name]]

#l <- func.predict.resp(item_prob, persons)

#func.predict.resp.all.items(df.cdm.structure.frac20, table.list, persons.with.id)
```

# simulate data
```{r}
Q <- frac20.q
N <- nrow(frac20$dat)

J <- nrow(Q)
gs <- coef(mod.dina, "gs")
```


```{r}
sim <- simGDINA(N,Q,gs.parm = gs,model = "DINA",gs.args=list(mono.constraint=TRUE))

dat <- extract(sim,what = "dat")

df.sim.resp.long <- as.data.frame(dat) %>% 
  add_column(id = seq(1, 536, 1)) %>% 
  pivot_longer(cols = -id,               # Specify all columns except pid to pivot
               names_to = "item",         # New column name for item names
               values_to = "resp") %>% 
  mutate(item = as.numeric(substring(item, 2)))
  
```

```{r}
df.sim.resp <- df.sim.resp.long %>% 
  left_join(Q)
```

# Out-of-sample prediction
for a given data
```{r}
set.seed(10)
```
```{r}
func.run.analysis <- function(df, cdm_model, heldout_rate, limited_skills){
  n_items = n_distinct(df$item)
  df.cdm.structure <- func.convert.Q.data.structure.with.held.out.limited.skills(df, heldout_rate, limited_skills)
  #df.resp.without.id <- df.cdm.structure$dat %>% arrange(id) %>% select(-id)
  mod.dina <- GDINA(df.cdm.structure$dat,df.cdm.structure$Q,model = cdm_model)
  
  
  # CDM model
  df.cdm.structure.q <- df.cdm.structure$Q %>% 
    mutate(item = seq(1, n_items, 1))
  
  q_matrix_long <-  df.cdm.structure.q %>%
    pivot_longer(cols = starts_with("Qmatrix"),  # Select columns that represent attributes
               names_to = "Attribute",  # Name for the new 'attribute' column
               values_to = "Value") 
  
  persons.with.id <- as.data.frame(personparm(mod.dina)) %>% 
    mutate(id = df.cdm.structure$id)
  
  coef <- coef(mod.dina, "itemprob")
  
  table.list <- func.get.table(n_items, coef, q_matrix_long)

  
  predict.matrix <- func.predict.resp.all.items(df.cdm.structure, table.list, persons.with.id, table.list)
  
  result <- predict.matrix  %>% 
    mutate(id = row_number()) %>% 
    pivot_longer(cols = starts_with("V"), names_to = "item", values_to = "prediction.cdm") %>% 
    left_join(df.cdm.structure$heldout)
 
  # 1PL model
  mod.Rasch <- mirt(df.cdm.structure$dat, 1, itemtype = "Rasch")
  
  thetas <- as.data.frame(fscores(mod.Rasch))
  
  params <- data.frame(coef(mod.Rasch, IRTpars = TRUE, simplify = TRUE))
  
  
  result.2 <- tibble(th = thetas$F1, id = df.cdm.structure$id) %>%
    left_join(df.cdm.structure$heldout) %>% 
    left_join(rownames_to_column(params, var = "item")) %>% 
    mutate(prediction.1PL = prob.irt(items.a, items.b, th)) %>% 
     dplyr ::select(id, item_num, item, resp, held_out, prediction.1PL)
  
  # 2PL model
  # mod.2PL <- mirt(df.cdm.structure$dat, 1, itemtype = "2PL")
  # 
  # thetas <- as.data.frame(fscores(mod.2PL))
  # 
  # params <- data.frame(coef(mod.2PL, IRTpars = TRUE, simplify = TRUE))
  # 
  # result.3 <- tibble(th = thetas$F1, id = df.cdm.structure$id) %>%
  #   left_join(df.cdm.structure$heldout) %>% 
  #   left_join(rownames_to_column(params, var = "item")) %>% 
  #   mutate(prediction.2PL = prob.irt(items.a, items.b, th)) %>% 
  #   select(id, item_num, item, resp, held_out, prediction.2PL)
  
  result.compared <- result %>% 
    left_join(result.2) 
  
  df.test <- result.compared %>% filter(held_out == 1) %>% 
    filter(!is.na(prediction.1PL), !is.na(prediction.cdm))
  
  imv <- imv.binary(df.test$resp, df.test$prediction.1PL, df.test$prediction.cdm)
  print(imv)
  return (df.test)
  #return (imv)
}
```

```{r}
df <- q_data$frac20$to_data_frame()
```

```{r}
#list.result <- c()
for (i in seq(1, 5, 1)){
  result <- func.run.analysis(df, "DINA", 0.2, 9)
  #list.result <- c(list.result, result)
}
```


```{r}
# list.result <- c()
# for (i in seq(1, 5, 1)){
#   result <- func.run.analysis(df.sim.resp, "DINA", 0.2)
#   list.result <- c(list.result, result)
# }
# mean(list.result)
# sd(list.result)
```
```{r}
for (i in seq(1, 5, 1)){
  result <- func.run.analysis(df.sim.resp, "DINA", 0.2, 9)
  #list.result <- c(list.result, result)
}
```

```{r}
#write.csv(result, "output/frac20_prediction.csv")
```

# Simulation Study
we have theta~N(0,Sigma) where Sigma is a 2x2 covariance matrix with variances=1 and correlation=r.

for theta 1, we have skills 1a and 1b. for a person, Pr(1a=1)=invlogit(theta_1-b_1a) where we can makes b_1a<b_1b to have a loewr and higher level skill

similarly for theta_2.

my claim is that when r~1, IMV(1pl,CDM) should be <=0. in that case, the 1PL should be able to do a good job of proxying for the CDM even when the CDM guess/slip probabilities govern responses.
(edited)

when we have r=0 or r<0, that is when IMV(1pl,CDM)>0 because now the structure is skills is not going to be something that a unidmensional theta model can approximate.
```{r}
logit<-function(x,a=1.7) 1/(1+exp(-a*x))

func.create.response <- function(b, th, np, ni) {
  th.mat<-matrix(th,np,ni,byrow=FALSE) 
  b.mat <- matrix(rep(b, np), nrow = np, byrow = TRUE)
  pr<-inv_logit(th.mat-b.mat)

  resp <- pr

  for (i in 1:ncol(resp)) {
    resp[,i]<-rbinom(nrow(resp),1,resp[,i])
  }

  return (data.frame(resp))
}
```

```{r}
df.create.sim.resp <- function(r, np){
  # Define the covariance matrix (Sigma)
  Sigma <- matrix(c(1, r, r, 1), nrow = 2, ncol = 2)
  theta_samples <- mvrnorm(n = np, mu = c(0, 0), Sigma = Sigma)
  
  df.thetas <- as.data.frame(theta_samples)
  
  plot(df.thetas$V1, df.thetas$V2)
  
  ni<-4
  
  b1.sample <- c(-2, 0, 0, 2)
  
  b2.sample <- c(-2, 0, 0, 2)
  
  th1.sample <- df.thetas$V1
  resp1.sample <- func.create.response(b1.sample, th1.sample, np, ni)
  th2.sample <- df.thetas$V2
  resp2.sample <- func.create.response(b2.sample, th2.sample, np, ni)
  
  new_names <- c("item_1", "item_2", "item_3", "item_4", "item_5", "item_6", "item_7", "item_8")
  
  df.designed.q <- tibble(item = new_names, Qmatrix__1 = c(1, 1, 1, 1, 0, 0, 0, 0), Qmatrix__2 = c(0, 1, 1, 1,  0, 0, 0, 0), Qmatrix__3  = c(0, 0, 0, 1, 0, 0,0, 0), Qmatrix__4 = c(0, 0, 0, 0, 1, 1, 1, 1), Qmatrix__5 = c(0, 0, 0, 0, 0, 1, 1, 1), Qmatrix__6 = c(0, 0, 0, 0, 0, 0, 0, 1)) 
  
  df.sim <- resp2.sample %>% 
    cbind(resp2.sample) %>% 
    set_names(new_names) %>% 
    add_column(id = seq(1, np, 1)) %>% 
    pivot_longer(cols = c(item_1, item_2, item_3, item_4, item_5, item_6, item_7, item_8), names_to = "item", values_to = "resp") %>%
    left_join(df.designed.q) %>% 
    relocate(item, id)
  
  return (df.sim)
  
}
```


## when r == 1
```{r}
#list.result <- c()
df.sim <- df.create.sim.resp(1, 1000)
for (i in seq(1, 10, 1)){
  result <- func.run.analysis(df.sim, "DINA", 0.2, 3)
  #list.result <- c(list.result, result)
}
#mean(list.result)
#sd(list.result)
```


## when r == 0
```{r}
#list.result <- c()
df.sim <- df.create.sim.resp(0, 1000)
for (i in seq(1, 10, 1)){
  result <- func.run.analysis(df.sim, "DINA", 0.2, 3)
  #list.result <- c(list.result, result)
}
#mean(list.result)
#sd(list.result)
```


# Ben's method

```{r}
N<-1000
r<- -0.5
##theta
th<-mvrnorm(N,mu=rep(0,2),Sigma=matrix(c(1,r,r,1),2,2))

##skills
logit<-function(x,a=1.7) 1/(1+exp(-a*x))
s1<-rbinom(N,1,logit(th[,1]+1))
s2<-rbinom(N,1,logit(th[,1]-1))
s3<-rbinom(N,1,logit(th[,2]+1))
s4<-rbinom(N,1,logit(th[,2]-1))
sk<-cbind(s1,s2,s3,s4)

##qmatrix
qm<-matrix(
    c(1,0,0,0,
      1,1,0,0,
      0,0,1,0,
      0,0,1,1
      ),4,4,byrow=TRUE)

##response probabilities for with g=s=0.1
g<-s<-0.1
    
resp<-list()
for (i in 1:nrow(qm)) {
    ii<-which(qm[i,]==1)
    z<-sk[,ii,drop=FALSE]
    rm<-rowMeans(z)
    p<-ifelse(rm==1,1-s,g)
    resp[[i]]<-rbinom(N,1,p)
}
resp<-do.call("cbind",resp)

```

```{r}
new_names <- c("item_1", "item_2", "item_3", "item_4")

skill_names <- c("Qmatrix__1", "Qmatrix__2", "Qmatrix__3", "Qmatrix__4")

df.sim <- as.data.frame(resp) %>% 
  set_names(new_names) %>% 
  add_column(id = seq(1, N, 1)) %>% 
  pivot_longer(cols = c(item_1, item_2, item_3, item_4), names_to = "item", values_to = "resp") %>% 
  left_join(as.data.frame(qm) %>% 
  set_names(skill_names) %>% 
  add_column(item = new_names))%>% 
  relocate(item, id)
```




```{r}
#list.result <- c()
for (i in seq(1, 4, 1)){
  result <- func.run.analysis(df.sim, "DINA", 0.2, 2)
  #list.result <- c(list.result, result)
}
#mean(list.result)
#sd(list.result)
```
```{r}
#list.result <- c()
for (i in seq(1, 4, 1)){
  result <- func.run.analysis(df.sim, "DINA", 0.2, 2)
  #list.result <- c(list.result, result)
}
#mean(list.result)
#sd(list.result)
```
```{r}
result
```

