---
title: "IRW explore"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    
---

# Data Preprocessing
## Load Packages
Let's first load the packages that we need for this chapter. 

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
library(dplyr)
library(mirt)
library('ggpubr')
library(Metrics)
library(hash)
library(ggplot2)
library(GGally)
library(gridExtra)
library(rstudioapi)
library(TAM)
library(mgcv)
library(GDINA)
```

## Load IRW
```{r}
#reliability 
dataset <- redivis::user("datapages")$dataset("item_response_warehouse",version='v7.0')

dataset_tables <- dataset$list_tables()
```


## Settings 
```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")

# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

```{r}
df <- dataset$table("frac20")$to_data_frame()
```

## Get CDM data
```{r}
##
print(length(dataset_tables))
names<-sapply(dataset_tables,function(x) x$name)
ii<-grep("metadata",names)
names(dataset_tables)<-names
if (length(ii)>0) dataset_tables<-dataset_tables[-ii]   
##
f<-function(table) table$list_variables()
nms<-lapply(dataset_tables,f)

##
f<-function(x) {
    nm<-sapply(x,function(x) x$name)
    any(grepl("^Qmat",nm))
}
test<-sapply(nms,f)
q_data<-dataset_tables[test]
print(length(q_data))
```
```{r}
df.frac20 <- q_data$frac20$to_data_frame()
```
# helper functions 

```{r}
imv.binary<-function(y, #outcomes
                     p1,#baseline
                     p2 #enhanced
                     ) {
    ##
    ll<-function(x,p) {
        z<-log(p)*x+log(1-p)*(1-x)
        z<-sum(z)/length(x)
        exp(z)
    }    
    loglik1<-ll(y,p1)
    loglik2<-ll(y,p2)
    getcoins<-function(a) {
        f<-function(p,a) abs(p*log(p)+(1-p)*log(1-p)-log(a))
        nlminb(.5,f,lower=0.001,upper=.999,a=a)$par
    }
    c1<-getcoins(loglik1)
    c2<-getcoins(loglik2)
    ew<-function(p1,p0) (p1-p0)/p0
    imv<-ew(c2,c1)
    imv
}
```

```{r}
prob.irt <-function(a, b,theta) {
  1 / (1 + exp(-a * (theta - b)))
  
}
```

# Convert to Q matrix and data 
```{r}
func.convert.Q.data.structure <- function(df){
  Q <- df %>% 
    select(-c(id, resp)) %>% 
    unique() %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
    arrange(item_num) %>% 
    select(-c(item, item_num))
  
  dat <- df %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
    select(item_num, id, resp) %>% 
    arrange(item_num) %>% 
    mutate(item = paste0("V",item_num)) %>% 
    select(-item_num) %>% 
    pivot_wider(names_from = item, values_from = resp) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- dat
  
  return (tables_list)
}
frac20 <- func.convert.Q.data.structure(df.frac20)
```

```{r}
func.convert.Q.data.structure.with.held.out <- function(df){
  Q <- df %>% 
    select(-c(id, resp)) %>% 
    unique() %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
    arrange(item_num) %>% 
    select(-c(item, item_num))
  
  df.held.out <- df %>% 
    mutate(item_num = str_extract(item, "\\d+") %>% as.numeric()) %>% 
    select(item_num, id, resp) %>% 
    arrange(item_num) %>% 
    mutate(item = paste0("V",item_num)) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = 0.2)) 
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
    select(-c(item_num, held_out)) %>% 
    pivot_wider(names_from = item, values_from = resp) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- df.dat
  tables_list[['heldout']] <- df.held.out
  
  return (tables_list)
}
```

# next steps: 
plan: try other models 
simulation data: simulation data from DINA 
use frac20 Q matrix, simulate new data 
simulate data from profile 
under situations 
skill heirachy 


# Analysis (All data)
fit DINA model

```{r}
mod.dina <- GDINA(frac20$dat,frac20$Q,model = "DINA")
```

```{r}
coef <- coef(mod.dina, "itemprob")

```

```{r}
frac20.q <- frac20$Q %>% 
  mutate(item = seq(1, 20, 1))

q_matrix_long <-  frac20.q %>%
  pivot_longer(cols = starts_with("Qmatrix"),  # Select columns that represent attributes
               names_to = "Attribute",  # Name for the new 'attribute' column
               values_to = "Value") 
```

get person estimate
```{r}
persons <- as.data.frame(personparm(mod.dina))

#colnames(persons) <- sub("^A", "V", colnames(persons))
```

```{r}
name <- paste0("Item", " ", 1)
df.coef <- as.data.frame(coef[[name]])
rownames(df.coef)
```


```{r}
func.get.table <- function(num, coef, q_matrix_long) {
  # Load necessary library
  library(dplyr)
  
  # Initialize an empty list to store the data frames
  tables_list <- list()
  
  for (i in seq(1, num, 1)) {
    
    # Filter the required attributes for the current item from q_matrix_long
    required_attribute <- q_matrix_long %>%
      filter(Value == 1, item == i) %>%
      arrange(Attribute)
    
    # Get the attribute names
    attributes <- required_attribute$Attribute
    
    # Create a dynamic name for the current item (e.g., "Item 1", "Item 2", etc.)
    name <- paste0("Item", " ", i)
    
    # Extract the coefficients for the current item and convert to data frame
    df.coef <- as.data.frame(coef[[name]])
    
    # Get the rownames of the coefficient data frame (the patterns)
    list.combination <- rownames(df.coef)
    
    # Function to convert pattern strings (like "P(000)", "P(100)", etc.)
    convert_pattern <- function(pattern) {
      # Remove "P(" and ")" to isolate the binary part
      binary_string <- gsub("[P()]", "", pattern)
      
      # Split the binary string into individual characters
      binary_vector <- as.numeric(unlist(strsplit(binary_string, "")))
      
      # Return the binary vector
      return(binary_vector)
    }
    
    # Extract probabilities and ensure they are numeric
    possibilities <- as.numeric(df.coef[[1]])
    
    # Convert the list.combination patterns to a matrix form
    attribute_matrix <- do.call(rbind, lapply(list.combination, convert_pattern))
    
    # Convert attribute_matrix to a data frame and ensure numeric
    df <- as.data.frame(attribute_matrix, stringsAsFactors = FALSE)
    df[] <- lapply(df, as.numeric)  # Coerce all columns to numeric
    
    # Add the probabilities as a separate numeric column
    df$Probability <- possibilities
    
    # Assign proper column names
    colnames(df) <- c(attributes, "Probability")
    colnames(df) <- sub("^Qmatrix__", "A", colnames(df))
    
    # Store the resulting data frame in the list
    tables_list[[name]] <- df
    
  }
  
  # Return the list of data frames
  return(tables_list)
}
table.list <- func.get.table(20, coef, q_matrix_long)
```


```{r}
func.predict.resp <- function(item_prob, persons){
  persons.prediction <- persons %>% 
    left_join(item_prob) 
  return(persons.prediction$Probability)
}

# want to return a dataframe with column i with "name: Vi"and values shall be the list l
func.predict.resp.all.items <- function(frac, item, persons){

  
  # Initialize an empty list to store predictions
  response_list <- list()
  
  n <- length(frac$dat) -1
  for (i in seq(1, n, 1)) {
    name <- paste0("Item", " ", i)
    
    item_prob <- table.list[[name]]
    
    l <- func.predict.resp(item_prob, persons)
    
    col_name <- paste0("V", i)
    response_list[[col_name]] <- l
  }
  response_df <- as.data.frame(response_list, stringsAsFactors = FALSE)
  
  # Optionally, add identifiers from 'persons' if needed
  # For example, if 'persons' has an ID column:
  # response_df$PersonID <- persons$ID
  
  return(response_df)
}

predict.matrix <- func.predict.resp.all.items(frac20, table.list, persons)
```
# Out-of-sample prediction
for a given data
```{r}
set.seed(10)
```
```{r}
func.run.analysis <- function(df){
  df.cdm.structure <- func.convert.Q.data.structure.with.held.out(df)
  df.resp.with.id <- df.cdm.structure$dat %>% arrange(id) %>% select(-id)
  mod.dina <- GDINA(df.resp.with.id,df.cdm.structure$Q,model = "DINA")
  
  # CDM model
  df.cdm.structure.q <- df.cdm.structure$Q %>% 
    mutate(item = seq(1, 20, 1))
  
  q_matrix_long <-  df.cdm.structure.q %>%
    pivot_longer(cols = starts_with("Qmatrix"),  # Select columns that represent attributes
               names_to = "Attribute",  # Name for the new 'attribute' column
               values_to = "Value") 
  
  persons <- as.data.frame(personparm(mod.dina))
  
  coef <- coef(mod.dina, "itemprob")
  
  table.list <- func.get.table(20, coef, q_matrix_long)
  
  predict.matrix <- func.predict.resp.all.items(df.cdm.structure, table.list, persons)
  
  result <- predict.matrix  %>% 
    mutate(id = row_number()) %>% 
    pivot_longer(cols = starts_with("V"), names_to = "item", values_to = "prediction.cdm") %>% 
    left_join(df.cdm.structure$heldout)
  
  # 1PL model
  mod.Rasch <- mirt(df.cdm.structure$dat %>% select(-id), 1, itemtype = "Rasch")
  
  thetas <- as.data.frame(fscores(mod.Rasch))
  
  params <- data.frame(coef(mod.Rasch, IRTpars = TRUE, simplify = TRUE))
  
  result.2 <- df.cdm.structure$dat %>% 
    add_column(th = thetas$F1) %>% 
    select(id, th) %>% 
    left_join(df.cdm.structure$heldout) %>% 
    left_join(rownames_to_column(params, var = "item")) %>% 
    mutate(prediction.1PL = prob.irt(items.a, items.b, th)) %>% 
    select(id, item_num, item, resp, held_out, prediction.1PL)
  
  result.compared <- result %>% left_join(result.2)
  
  df.test <- result.compared %>% 
    filter(held_out == 1)
  
  print(imv.binary(df.test$resp, df.test$prediction.1PL, df.test$prediction.cdm))
  
  return (result.compared)
}
```
```{r}
df.frac20 <- q_data$frac20$to_data_frame()
```

```{r}
for (i in seq(1, 5, 1)){
  print(i)
  result <- func.run.analysis(df.frac20)
}
```
```{r}
result %>% 
  filter(prediction.cdm == 0)
```

```{r}
ggplot(result %>% filter(item == "V3"), mapping = aes(x = prediction.1PL,
                     y = prediction.cdm)) + geom_smooth(method = "lm")  + stat_cor() + geom_point() 
```

```{r}
df.test <- result %>% 
  filter(held_out == 1)
```

```{r}
imv.binary(df.test$resp, df.test$prediction.1PL, df.test$prediction.cdm)
```


```{r}
#write.csv(result, "output/frac20_prediction.csv")
```
