---
title: "IRW explore"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    
---

# Data Preprocessing
## Load Packages
Let's first load the packages that we need for this chapter. 

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
library(dplyr)
library(mirt)
library('ggpubr')
library(Metrics)
library(hash)
library(ggplot2)
library(GGally)
library(gridExtra)
library(rstudioapi)
library(TAM)
library(mgcv)
library(GDINA)
library(MASS)
library(imv)
```
```{r}
source("bd/00funs.R")
```

```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")

# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

# Load IRW
```{r}
#reliability 
dataset <- redivis::user("datapages")$dataset("item_response_warehouse",version='v7.0')

dataset_tables <- dataset$list_tables()
```
```{r}
load("data/frac20test.RData")
```
```{r}
df.frac20.test <- df
```

```{r}
fit1 <- GDINA(dat = frac20$dat, Q = frac20$Q, model = "DINA")
coef(object = fit1, what = "delta")

as.data.frame(coef(object = fit1, what = "gs"))
```

## Get CDM data
```{r}
##
print(length(dataset_tables))
names<-sapply(dataset_tables,function(x) x$name)
ii<-grep("metadata",names)
names(dataset_tables)<-names
if (length(ii)>0) dataset_tables<-dataset_tables[-ii]   
##
f<-function(table) table$list_variables()
nms<-lapply(dataset_tables,f)

##
f<-function(x) {
    nm<-sapply(x,function(x) x$name)
    any(grepl("^Qmat",nm))
}
test<-sapply(nms,f)
q_data<-dataset_tables[test]
print(length(q_data))
```


# helper functions 
## IRT
```{r}
prob.irt <-function(a, b,theta) {
  1 / (1 + exp(-a * (theta - b)))
}
```

## Convert to Q matrix and data 

```{r}
func.convert.Q.data.structure.with.held.out <- function(df, heldout_rate){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  df.held.out <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = heldout_rate)) %>% 
     dplyr ::arrange(id)
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
     dplyr ::select(-c(held_out)) %>% 
    pivot_wider(names_from = item, values_from = resp) %>% 
     dplyr ::arrange(id) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- df.dat %>%  dplyr ::select(-id)
  tables_list[['heldout']] <- df.held.out
  tables_list[['id']] <- df.dat$id
  
  return (tables_list)
}
```

```{r}
func.convert.Q.data.structure.with.k.fold <- function(df, kfold){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  Q <- Q[, order(colnames(Q))]
  
  assignments <- sample(rep(1:kfold, length.out = length(df$id)))
  
  df.kfold <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    add_column(gr = assignments) %>% 
     dplyr ::arrange(id)
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat_kfold']] <- df.kfold
  
  return (tables_list)
}

```


## OOS prediction
```{r}
# func.run.analysis <- function(df, cdm_model, if_oos, heldout_rate, cdm_pr_estimate = cdm.pr){
#   n_items = n_distinct(df$item)
#   item_list <- sort(unique(df$item))
#   
#   df.cdm.structure <- func.convert.Q.data.structure.with.held.out(df, heldout_rate)
#   
#   predict.matrix <- cdm_pr_estimate(df.cdm.structure$dat,df.cdm.structure$Q, model = cdm_model)
#   
#   result <- as.data.frame(predict.matrix) %>% 
#     set_names(item_list) %>% 
#     mutate(id = row_number()) %>% 
#     pivot_longer(cols = item_list, names_to = "item", values_to = "prediction.cdm") %>% 
#     left_join(df.cdm.structure$heldout)
#  
#   # IRT model
#   mod.irt <- mirt(df.cdm.structure$dat, 1, itemtype = "2PL")
#   
#   thetas <- as.data.frame(fscores(mod.irt))
#   
#   params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
#   
#   result.2 <- tibble(th = thetas$F1, id = df.cdm.structure$id) %>%
#     left_join(df.cdm.structure$heldout) %>% 
#     left_join(rownames_to_column(params, var = "item")) %>% 
#     mutate(prediction.irt = prob.irt(items.a, items.b, th)) %>% 
#      dplyr ::select(id, item, resp, held_out, prediction.irt)
#   
#   result.compared <- result %>% 
#     left_join(result.2) 
#   
#   
#   
#   if (if_oos) {
#     df.test <- result.compared %>% 
#     filter(held_out == 1) %>% 
#     filter(!is.na(prediction.irt), !is.na(prediction.cdm))
#   } else {
#     df.test <- result.compared %>% 
#     filter(!is.na(prediction.irt), !is.na(prediction.cdm))
#   }
#   
#   imv <- imv::imv.binary(df.test$resp, df.test$prediction.irt, df.test$prediction.cdm)
#   print(imv)
#   return (imv)
# }
```


```{r}
func.run.analysis.kfold <- function(df, cdm_model = "DINA", irt_model = "2PL", if_oos = TRUE, kfold = 5, cdm_pr_estimate = cdm.pr.marg){
  n_items = n_distinct(df$item)
  item_list <- sort(unique(df$item))
  
  df.cdm.structure <- func.convert.Q.data.structure.with.k.fold(df, kfold = kfold)
  
  func.get.prediction.result <- function(df.tmp, dat, qm){
    # CDM model
    df.dat <- dat %>% 
      dplyr ::select(-gr) %>% 
      dplyr :: arrange(item) %>% 
      pivot_wider(names_from = item, values_from = resp) %>% 
      dplyr ::arrange(id) 
    
    df.dat.no.id <- df.dat %>% dplyr :: select(-id)
    
    predict.matrix <- cdm_pr_estimate(df.dat.no.id, qm, model = cdm_model)
    
    result <- as.data.frame(predict.matrix) %>% 
      set_names(item_list) %>% 
      mutate(id = row_number()) %>% 
      pivot_longer(cols = item_list, names_to = "item", values_to = "prediction.cdm") %>% 
      left_join(df.tmp)
    
    # IRT model
    mod.irt <- mirt(df.dat.no.id, 1, itemtype = irt_model)
    
    thetas <- as.data.frame(fscores(mod.irt))

    params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
    
    result.2 <- tibble(th = thetas$F1, id = df.dat$id) %>%
      left_join(df.tmp) %>% 
      left_join(rownames_to_column(params, var = "item")) %>% 
      mutate(prediction.irt = prob.irt(items.a, items.b, th)) %>% 
      dplyr ::select(id, item, resp, gr, prediction.irt)
    
    result.compared <- result %>% 
      left_join(result.2) 
  }
  
  list.imv <- list()
  
  list.cor <- list()
  
  if (!if_oos) {
    # full data set
    result.compared <- func.get.prediction.result(df.cdm.structure$dat_kfold, df.cdm.structure$dat_kfold, df.cdm.structure$Q)
    
    df.compare <- result.compared %>% 
        filter(!is.na(prediction.irt), !is.na(prediction.cdm))
      
    imv <- imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm)
    
    cor <- cor(df.compare$prediction.irt, df.compare$prediction.cdm)
      
    list.imv <- c(list.imv, imv) 
    
    list.cor <- c(list.cor, cor)
    
  }
  else {
    for (i in seq(1, kfold, 1)) {
      df.tmp <- df.cdm.structure$dat_kfold
      
      df.train <- df.tmp %>% 
        filter(gr != i) 
      
      result.compared <- func.get.prediction.result(df.tmp, df.train, df.cdm.structure$Q)
      
      df.compare <- result.compared %>% 
        filter(gr == i) %>% 
        filter(!is.na(prediction.irt), !is.na(prediction.cdm))
      
      imv <- imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm)
      
      cor <- cor(df.compare$prediction.irt, df.compare$prediction.cdm)
      
      list.imv <- c(list.imv, imv) 
      
      list.cor <- c(list.cor, cor)
    }
  }
  
  imv_reuslt <- list()
  imv_reuslt[['imv_mean']] <- mean(unlist(list.imv))
  imv_reuslt[['imv_sd']] <- sd(unlist(list.imv))
  imv_reuslt[['cor_mean']] <- mean(unlist(list.cor))
  imv_reuslt[['cor_sd']] <- sd(unlist(list.cor))
  
  return(imv_reuslt)
}
```


# Empirical analysis

## sanity check 

### Q1: check if q_matrix is correct?

```{r}
frac20_irw <- q_data$frac20$to_data_frame() 

item_list <- sort(unique(frac20_irw $item))

df.cdm.structure <- func.convert.Q.data.structure.with.held.out(frac20_irw, 0.2)
```

```{r}
hi <- func.run.analysis.debug(frac20_irw, "DINA", TRUE, cdm_pr_estimate = cdm.pr.marg)
```
```{r}
hi <- func.run.analysis.kfold(frac20_irw, "DINA", "2PL", TRUE, 5, cdm.pr.marg)
```
```{r}
func.run.analysis.kfold(frac20_irw, "DINA", "2PL", FALSE, 5, cdm.pr.marg)
```
### Q2: check if frac20 cdm prediction makes sense?

```{r}
predict.matrix <- cdm.pr(df.cdm.structure$dat,df.cdm.structure$Q, model = "DINA")
  
result <- as.data.frame(predict.matrix) %>% 
    set_names(item_list) %>% 
    mutate(id = row_number()) %>% 
    pivot_longer(cols = item_list, names_to = "item", values_to = "prediction.cdm") %>% 
    left_join(df.cdm.structure$heldout)
 
  # IRT model
mod.irt <- mirt(df.cdm.structure$dat, 1, itemtype = "2PL")
  
thetas <- as.data.frame(fscores(mod.irt))
  
params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
  
result.2 <- tibble(th = thetas$F1, id = df.cdm.structure$id) %>%
    left_join(df.cdm.structure$heldout) %>% 
    left_join(rownames_to_column(params, var = "item")) %>% 
    mutate(prediction.irt = prob.irt(items.a, items.b, th)) %>% 
     dplyr ::select(id, item, resp, held_out, prediction.irt)
  
result.compared <- result %>% 
    left_join(result.2) 
```

prediction correlation between cdm and irt
```{r}
cor(result.compared$prediction.cdm, result.compared$prediction.irt)
```
```{r}
hi <- result.compared %>% 
  filter(held_out == 0) 

imv::imv.binary(hi$resp, hi$prediction.irt, hi$prediction.cdm)
```

### Q3: check if any item?
```{r}
# df.item.check <- result.compared %>% 
#     mutate(prediction.cdm.binary = ifelse(runif(n()) < prediction.cdm, 1, 0),  # Roll dice based on prediction.cdm probability
#          prediction.irt.binary = ifelse(runif(n()) < prediction.irt, 1, 0)   # Roll dice based on prediction.irt probability
#          ) %>% 
#   mutate(cdm.correct = ifelse(resp == prediction.cdm.binary, 1, 0), 
#          irt.correct = ifelse(resp == prediction.irt.binary, 1, 0)) %>% 
#   group_by(item, held_out) %>% 
#   dplyr :: summarise(p.cdm.correct = mean(cdm.correct), p.irt.correct = mean(irt.correct)) %>% 
#   arrange(p.cdm.correct)
# 
# ggplot(df.item.check %>% mutate(held_out = as.character(held_out)),
#        mapping = aes(x = p.cdm.correct, 
#                      y = p.irt.correct, 
#                      color = held_out)) +
#   # Connect points with the same 'item'
#   geom_line(aes(group = item), color = "gray") +  # Lines for same 'item'
#   geom_text(aes(label = item)) +  # Text labels for items
#   scale_color_manual(values = c("black", "darkred")) + 
#   xlim(0.3, 1) + ylim(0.3, 1) + 
#   coord_equal() +
#   theme_minimal() + stat_cor()
```

## Run all data

```{r}
func.run.empirical.data <- function(cdm_model = "DINA", irt_model = "2PL", cdm_pr_estimate = cdm.pr) {
  # Initialize empty lists to store results
  all_results <- list()
  oos_results <- list()
  
  # Iterate over objects in q_data (adjust the loop as needed for your actual structure)
  for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa")) {
      print(obj_name)
      
      # Convert the object to a data frame
      df <- q_data[[obj_name]]$to_data_frame()
      
      # Initialize lists to store results for this object
      object_1_results <- list()
      object_2_results <- list()
      
      # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      # Run the first type of analysis once
      result <- func.run.analysis.kfold(df, cdm_model, irt_model, FALSE, 5, cdm.pr.marg)
      object_1_results[[i]] <- result
      
      # Run the second type of analysis multiple times
      result <- func.run.analysis.kfold(df, cdm_model, irt_model, TRUE, 5, cdm.pr.marg)
      object_2_results[[i]] <- result
    
      
      # Combine results for this object and store them in lists
      all_results[[obj_name]] <- do.call(rbind, object_1_results)
      oos_results[[obj_name]] <- do.call(rbind, object_2_results)
    }
  }
  
  # Combine all results into a single data frame
  final_all_results <- do.call(rbind, lapply(names(all_results), function(name) {
    cbind(object = name, all_results[[name]])
  }))
  
  final_oos_results <- do.call(rbind, lapply(names(oos_results), function(name) {
    cbind(object = name, oos_results[[name]])
  }))
  
  df.imv.result <- as.data.frame(final_all_results) %>% 
  dplyr :: select(object, imv_mean, cor_mean) %>% 
  set_names("dataname", "imv_all_response", "cor_all_response") %>% 
  mutate(dataname = as.character(dataname),
    imv_all_response = as.numeric(imv_all_response), 
         cor_all_response = as.numeric(cor_all_response)) %>% 
  left_join(as.data.frame(final_oos_results) %>% 
              set_names("dataname", "imv_kfold_response", "imv_kfold_response_sd", "cor_kfold_response", "cor_kfold_response_sd") %>% 
  mutate(dataname = as.character(dataname),
    imv_kfold_response = as.numeric(imv_kfold_response), 
         imv_kfold_response_sd = as.numeric(imv_kfold_response_sd), 
         cor_kfold_response = as.numeric(cor_kfold_response), 
         cor_kfold_response_sd = as.numeric(cor_kfold_response_sd)))
  
  # Return the comparison data frame
  return(df.imv.result)
}
```

## 2PL 
```{r}
set.seed(10)

df.imv.compare.1 <- func.run.empirical.data(cdm_model = "DINA", irt_model = "2PL")

df.imv.compare.2 <- func.run.empirical.data(cdm_model = "DINO", irt_model = "2PL")

df.imv.compare.3 <- func.run.empirical.data(cdm_model = "GDINA", irt_model = "2PL")
```

```{r}
df.imv.compare.result <- df.imv.compare.1 %>% 
  mutate(cdm_model = "DINA") %>% 
  rbind(df.imv.compare.2 %>% mutate(cdm_model = "DINO")) %>% 
  rbind(df.imv.compare.3 %>% mutate(cdm_model = "GDINA")) %>% 
  relocate(dataname, cdm_model) %>% 
  arrange(dataname) %>% 
  dplyr :: select(dataname, cdm_model, "imv_all_response", "imv_kfold_response", "imv_kfold_response_sd")
```


## 1PL 
```{r}
set.seed(10)

df.imv.compare.1.1pl <- func.run.empirical.data(cdm_model = "DINA", irt_model = "Rasch")

df.imv.compare.2.1pl <- func.run.empirical.data(cdm_model = "DINO", irt_model = "Rasch")

df.imv.compare.3.1pl <- func.run.empirical.data(cdm_model = "GDINA", irt_model = "Rasch")
```

# LaTeX table
```{r}
library(kableExtra)
```

```{r}
# Generate LaTeX Table with Styling
table_latex <- kable(df.imv.compare.result, format = "latex", caption = "IMV(2PL, CDM) from IRW empirical datasets") %>%
  kable_styling(latex_options = c("hold_position"))

# Print LaTeX Code
cat(table_latex)
```

