---
title: "Empirical analysis with IRW data"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    
---

# Data Preprocessing
## Load Packages
```{r}
#devtools::install_github("ben-domingue/imv", ref="main")
#remotes::install_github("hansorlee/irwpkg")
```

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
library(dplyr)
library(mirt)
library('ggpubr')
library(Metrics)
library(hash)
library(ggplot2)
library(GGally)
library(gridExtra)
library(rstudioapi)
library(TAM)
library(mgcv)
library(GDINA)
library(MASS)
library(imv)
library(irw)
library(pks) 
library(tidyr)

library(kableExtra)
library(conflicted)
conflict_prefer("extract", "tidyr")

```
```{r}
source("00funs.R")
```

```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")

# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```


# Load IRW
```{r}
dataset <- redivis::user("datapages")$dataset("item_response_warehouse",version='v10.0')

dataset_tables <- dataset$list_tables()
```

## Get CDM data
```{r}
##
print(length(dataset_tables))
names<-sapply(dataset_tables,function(x) x$name)
ii<-grep("metadata",names)
names(dataset_tables)<-names
if (length(ii)>0) dataset_tables<-dataset_tables[-ii]   
##
f<-function(table) table$list_variables()
nms<-lapply(dataset_tables,f)

##
f<-function(x) {
    nm<-sapply(x,function(x) x$name)
    any(grepl("^Qmat",nm))
}
test<-sapply(nms,f)
q_data<-dataset_tables[test]
print(length(q_data))
```


# helper functions 
## IRT
```{r}
prob.irt <-function(a, b,theta) {
  1 / (1 + exp(-a * (theta - b)))
}
```

## Convert to Q matrix and data 
```{r}
func.convert.Q.data.structure.with.held.out <- function(df, heldout_rate){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  df.held.out <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = heldout_rate)) %>% 
     dplyr ::arrange(id)
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
     dplyr ::select(-c(held_out)) %>% 
    pivot_wider(names_from = item, values_from = resp) %>% 
     dplyr ::arrange(id) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- df.dat %>%  dplyr ::select(-id)
  tables_list[['heldout']] <- df.held.out
  tables_list[['id']] <- df.dat$id
  
  return (tables_list)
}

func.convert.Q.data.structure.with.k.fold <- function(df, kfold){
  
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  Q <- Q[, colSums(Q != 0) > 0]
  
  Q <- Q[, order(colnames(Q))]
  
  assignments <- sample(rep(1:kfold, length.out = length(df$id)))
  
  df.kfold <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    add_column(gr = assignments) %>% 
     dplyr ::arrange(id)
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat_kfold']] <- df.kfold
  
  return (tables_list)
}

```


## OOS prediction

```{r}
func.run.analysis.kfold <- function(df, cdm_model = "DINA", irt_model = "2PL", if_oos = TRUE, kfold = 5, cdm_pr_estimate = cdm.pr.marg.update) {
  n_items <- n_distinct(df$item)
  item_list <- sort(unique(df$item))
  
  df.cdm.structure <- func.convert.Q.data.structure.with.k.fold(df, kfold = kfold)
  
  func.get.prediction.result <- function(df.tmp, dat, qm){

    df.ids <- df.tmp %>% group_by(id) %>%
      tally()

    df.dat <- dat %>%
      dplyr ::select(-gr) %>%
      dplyr :: arrange(item) %>%
      pivot_wider(names_from = item, values_from = resp) %>%
      dplyr ::arrange(id)

    df.dat.no.id <- df.dat %>% dplyr :: select(-id)

    # IRT model
    mod.irt <- mirt(df.dat.no.id, 1, itemtype = irt_model, technical = list(NCYCLES = 5000))
  
    if (!extract.mirt(mod.irt, "converged")) {
      print("irt not converged")
      return(NULL)
    }

    thetas <- as.data.frame(fscores(mod.irt))

    params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
    
    # CDM model
    predict.matrix.cdm.map <- cdm_pr_estimate(df.dat.no.id, qm, model = cdm_model, "MAP")
    if (is.null(predict.matrix.cdm.map)) {
      print("cdm map not converged")
      return(NULL)
    }
    
    predict.matrix.cdm.mp <- cdm_pr_estimate(df.dat.no.id, qm, model = cdm_model, "mp")
    if (is.null(predict.matrix.cdm.mp)) {
      print("cdm mp not converged")
      return(NULL)
    }

    result.cdm.map <- as.data.frame(predict.matrix.cdm.map) %>%
      set_names(item_list) %>%
      mutate(id = df.ids$id) %>%
      pivot_longer(cols = item_list, names_to = "item", values_to = "prediction.cdm.map")

    result.cdm.mp <- as.data.frame(predict.matrix.cdm.mp) %>%
      set_names(item_list) %>%
      mutate(id = df.ids$id) %>%
      pivot_longer(cols = item_list, names_to = "item", values_to = "prediction.cdm.mp")

    result.irt <- tibble(th = thetas$F1, id = df.dat$id) %>%
      left_join(df.tmp) %>%
      left_join(rownames_to_column(params, var = "item")) %>%
      mutate(prediction.irt = prob.irt(items.a, items.b, th)) %>%
      dplyr ::select(id, item, resp, gr, prediction.irt)

    result.compared <- result.cdm.map %>%
      left_join(result.cdm.mp) %>%
      left_join(df.tmp) %>%
      left_join(result.irt)
  }
  
  list.imv.1 <- list()
  list.imv.2 <- list()
  list.imv.3 <- list()

  if (!if_oos) {
    result.compared <- func.get.prediction.result(
      df.cdm.structure$dat_kfold, 
      df.cdm.structure$dat_kfold, 
      df.cdm.structure$Q
    )
    
    if (is.null(result.compared)) return(NA)
    
    df.compare <- result.compared %>% 
      filter(!is.na(prediction.irt), 
             !is.na(prediction.cdm.map), 
             !is.na(prediction.cdm.mp))
    
    list.imv.1 <- c(imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm.map))
    list.imv.2 <- c(imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm.mp))
    list.imv.3 <- c(imv::imv.binary(df.compare$resp, df.compare$prediction.cdm.map, df.compare$prediction.cdm.mp))
    
  } else {
    for (i in seq(1, kfold, 1)) {
      df.tmp <- df.cdm.structure$dat_kfold
      
      df.train <- df.tmp %>% 
        filter(gr != i) 
      
      result.compared <- func.get.prediction.result(df.tmp, df.train, df.cdm.structure$Q)
      
      if (is.null(result.compared)) return(next)
      
      df.compare <- result.compared %>% 
        filter(gr == i) %>% 
        filter(!is.na(prediction.irt), !is.na(prediction.cdm.map), !is.na(prediction.cdm.mp))
      
      imv.1 <- c(imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm.map))
      imv.2 <- c(imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm.mp))
      imv.3 <- c(imv::imv.binary(df.compare$resp, df.compare$prediction.cdm.map, df.compare$prediction.cdm.mp))
      
      list.imv.1 <- c(list.imv.1, imv.1) 
      list.imv.2 <- c(list.imv.2, imv.2) 
      list.imv.3 <- c(list.imv.3, imv.3) 
      
    }
  }

  return(c(
    mean(unlist(list.imv.1)), sd(unlist(list.imv.1)),
    mean(unlist(list.imv.2)), sd(unlist(list.imv.2)),
    mean(unlist(list.imv.3)), sd(unlist(list.imv.3))
  ))
}
```


# Empirical analysis

## Organize all data
```{r}
empirical_q_data
```

```{r}
func.run.empirical.data <- function(q_data, cdm_model = "DINA", irt_model = "2PL", cdm_pr_estimate = cdm.pr.marg.update) {
  # Initialize empty lists to store results
  all_results <- list()
  oos_results <- list()
  name_results <- c()
  
  # Iterate over objects in q_data (adjust the loop as needed for your actual structure)
  for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa", "cdm_hr")) {
      print(obj_name)
      
      # Convert the object to a data frame
      #df <- q_data[[obj_name]]$to_data_frame()
      
      if (is.data.frame(q_data[[obj_name]])) {
        df <- q_data[[obj_name]]
      } 
      else {
        df <- q_data[[obj_name]]$to_data_frame()  # Assuming the object has this method
      }
      
      # Initialize lists to store results for this object
      object_1_results <- list()
      object_2_results <- list()
      
      # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      if ("booklet" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "booklet"]
      }
      
      # Run the first type of analysis once
      object_1_results <- func.run.analysis.kfold(df, cdm_model, irt_model, FALSE, 5, cdm_pr_estimate)
      
      # Run the second type of analysis multiple times
      object_2_results <- func.run.analysis.kfold(df, cdm_model, irt_model, TRUE, 5, cdm_pr_estimate)
    
      # Combine results for this object and store them in lists
      all_results <- rbind(all_results, object_1_results)
      oos_results <- rbind(oos_results, object_2_results)
      name_results <- c(name_results, obj_name)
      
    }
  }
  
  colnames(all_results) <- c("mean_imv_1",  "sd_imv_1", "mean_imv_2",  "sd_imv_2", "mean_imv_3",  "sd_imv_3")
  colnames(oos_results) <- c("mean_imv_1",  "sd_imv_1", "mean_imv_2",  "sd_imv_2", "mean_imv_3",  "sd_imv_3")
  
  df.imv.result <- as.data.frame(all_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "all") %>% 
    rbind(as.data.frame(oos_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "kfold")) %>% 
    mutate(mean_imv_1 = as.numeric(mean_imv_1), 
         sd_imv_1 = as.numeric(sd_imv_1), 
         mean_imv_2 = as.numeric(mean_imv_2), 
         sd_imv_2 = as.numeric(sd_imv_2),
         mean_imv_3 = as.numeric(mean_imv_3), 
         sd_imv_3 = as.numeric(sd_imv_3))
  
  rownames(df.imv.result) <- NULL
    
  # Return the comparison data frame
  return(df.imv.result)
}
```


### irw-data
```{r}
set.seed(10)
```

### non-irw data
timss data 
```{r}
timss2011 <- readRDS("../data/TIMSS2011data.rds")

timss_data <-timss2011$data 

item_names <- colnames(timss_data)

timss_data_with_id <- rownames_to_column(timss_data, var = "id")

timss_q <- timss2011$Qc

timss_q_clean <- timss_q[-12, ][,-1][,-1]

timss2011_ready <- timss_data_with_id %>% 
  pivot_longer(cols = -id, names_to = "item", values_to = "resp") %>% 
  left_join(as.data.frame(timss_q_clean) %>% 
  add_column(item = item_names)) %>% 
  filter(resp!= 2)
  
```

mental health data
```{r}
load("../data/CDM_AlcoholApplication_Data_IRW.RData")

mental_health_ready <- df %>% 
  rename(resp = "response", V1 = "Qmatrix_skill1", V2 = "Qmatrix_skill2", V3 = "Qmatrix_skill3", V4 = "Qmatrix_skill4")
```

roar pa data
```{r}
load("../data/ROAR_PA.RData")

roar_pa_ready <- df %>% 
  dplyr ::select(-rt) %>% 
  mutate(id = as.character(id))
```


```{r}
convert.fraction.data <- function(dat, Q) {
  
  item_names <- colnames(dat)
  
  dat_with_id <- rownames_to_column(dat, var = "id")
  
  dat_ready <- dat_with_id %>%
    pivot_longer(cols = -id, names_to = "item", values_to = "resp") %>%
    left_join(as.data.frame(Q) %>% 
                add_column(item = item_names)) 
  return (dat_ready)
}
```

```{r}
q_data2 <- list(
  timss_2011 = timss2011_ready, 
  mental_health = mental_health_ready,
  roar_pa = roar_pa_ready,
  frac15q5 = convert.fraction.data(data.fraction1$data, data.fraction1$q.matrix),
  frac11q3 = convert.fraction.data(data.fraction2$data, data.fraction2$q.matrix1),
  frac11q5 = convert.fraction.data(data.fraction2$data, data.fraction2$q.matrix2)
)
```



```{r}
empirical_q_data <- append(q_data, q_data2)
```


## sanity check 

### Q1: check if q_matrix is correct?

get from irw
```{r}
df_frac20 <- q_data$frac20$to_data_frame()

df_frac20_clean <- df_frac20 %>% mutate(item = as.integer(substring(item, 6)))

df_frac20_cdm_structure <- func.convert.Q.data.structure.with.k.fold(df_frac20_clean, 0)
```



### Q2: check if frac20 cdm prediction makes sense?

```{r}
func.run.analysis.kfold(df_frac20, "GDINA", "2PL", FALSE, 5, cdm.pr.marg.update)
```

```{r}
func.run.analysis.kfold(df_frac20, "DINA", "2PL", TRUE, 5, cdm.pr.marg.update)
```


### Q3: check identifiability of all data 
```{r}

for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa", "cdm_hr")) {
      print(obj_name)
      df <- q_data[[obj_name]]$to_data_frame() 
      structure <- func.convert.Q.data.structure.with.k.fold(df, 0)
      print(check.qm.omplete(structure$Q))
    }
}

```
```{r}
for (obj_name in names(q_data2)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa", "cdm_hr")) {
      print(obj_name)
      df <- q_data2[[obj_name]]
      structure <- func.convert.Q.data.structure.with.k.fold(df, 0)
      print(check.qm.omplete(structure$Q))
    }
}
  
```
### Q4: cdm deep dive
```{r}
df <- GDINA::realdata_Tatsuoka1990$dat
Q <- GDINA::realdata_Tatsuoka1990$Q

#### GDINA estimation
gest <- GDINA(df,Q,mono.constraint = TRUE,control = list(nstarts=200))
### Q-matrix validation
qv <- Qval(gest,method = "wald")
qv
#### Fit GDINA with validated Q
gest2 <- GDINA(df,qv$sug.Q,mono.constraint = TRUE,control = list(nstarts=200))

#### find reduce models
y <- modelcomp(gest2,method = "Wald")
y

#######fit the reduced models using the validated Q
rest2 <- GDINA(df,qv$sug.Q,model=y$selected.model$models,mono.constraint = TRUE,control = list(nstarts=200))

##### compare different models
anova(gest,gest2,rest2)
```
```{r}
df_1 <- convert.fraction.data(df, Q)
func.run.analysis.kfold(df_1, "GDINA", "2PL", TRUE, 5, cdm.pr.marg.update)
```
```{r}
df_2 <- convert.fraction.data(df, qv$sug.Q)
func.run.analysis.kfold(df_2, "GDINA", "2PL", TRUE, 5, cdm.pr.marg.update)
```

```{r}
df_2 <- convert.fraction.data(df, qv$sug.Q)
func.run.analysis.kfold(df_2, y$selected.model$models, "2PL", TRUE, 5, cdm.pr.marg.update)
```

### Q5 new data 
Ma & Jiang 2021

```{r}
data(probability)
pb <- probability[!is.na(probability$b201), sprintf("b2%.2i", 1:12)]
Q1=sf1 <- read.table(header = TRUE, text = "
                    item cp id pb un
                    1  0  0  1  0
                    2  1  0  0  0
                    3  0  0  0  1
                    4  0  1  0  0
                    5  1  0  1  0
                    6  1  0  1  0
                    7  0  0  1  1
                    8  0  0  1  1
                    9  0  1  1  0
                    10  1  1  0  0
                    11  1  1  1  0
                    12  0  1  1  1
                    ")
Q1=Q1[,-1]

colnames(Q1) <- paste0("A", 1:4)
colnames(pb) <- paste0("V", 1:12)

pb_ready <- convert.fraction.data(pb, Q1)

```

```{r}
func.run.analysis.kfold(pb_ready, "GDINA", "2PL", TRUE, 5, cdm.pr.marg.update)
```

Tang & Zhan 2021
```{r}
Zhan_T1 <- read_csv("../data/zhan_t1/Zhan_T1.csv")

Zhan_T1_Q <- read.table("../data/zhan_t1/Q.txt")


colnames(Zhan_T1_Q) <- paste0("A", 1:6)
```
```{r}
rational_number_ready <- convert.fraction.data(Zhan_T1, Zhan_T1_Q)
```
```{r}
func.run.analysis.kfold(rational_number_ready, "DINA", "2PL", TRUE, 5, cdm.pr.marg.update)
```

```{r}
func.run.analysis.kfold(rational_number_ready, "GDINA", "2PL", TRUE, 5, cdm.pr.marg.update)
```


## Run all data
```{r}
set.seed(10)
df.imv.compare.1 <- func.run.empirical.data(empirical_q_data, cdm_model = "DINA", irt_model = "2PL")
```

```{r}
df.imv.compare.2 <- func.run.empirical.data(empirical_q_data, cdm_model = "GDINA", irt_model = "2PL")
```

## Analysis
```{r}
df.imv.compare.combined.raw <- df.imv.compare.1 %>% 
  mutate(cdm_model = "DINA") %>% 
  rbind(df.imv.compare.2 %>% mutate(cdm_model = "GDINA")) %>% 
  relocate(dataname, cdm_model, sample_method)

#write_csv(df.imv.compare.combined.raw, "../empirical_output/empirical_output_imv_raw.csv")
```

### IMV(MAP, mp)
```{r}
df.imv.estimator <- df.imv.compare.combined.raw %>% 
  dplyr :: select(-c(mean_imv_1, mean_imv_2, sd_imv_1, sd_imv_2)) %>% 
  pivot_wider(names_from = sample_method, values_from = c(mean_imv_3,sd_imv_3) ) %>% 
  arrange(dataname, cdm_model) %>% 
  relocate(dataname, cdm_model) %>% 
  dplyr ::select(-sd_imv_3_all) 
```

```{r}
# Generate LaTeX Table with Styling
table_latex <- kable(df.imv.estimator , format = "latex", caption = "IMV(MAP, mp) from IRW empirical datasets") %>%
  kable_styling(latex_options = c("hold_position"))

df.sorted <- df.imv.estimator %>% 
  arrange(dataname) %>%
  mutate(across(where(is.numeric), round, 5)) %>%
mutate(dataname = ifelse(duplicated(dataname), "", dataname))

# Create LaTeX table with midrules between dataname groups
table_latex <- kable(
  df.sorted, 
  format = "latex", 
  booktabs = TRUE,  # Ensures \toprule, \midrule, \bottomrule
  caption = "IMV(2PL, CDM) from IRW empirical datasets",
  col.names = c("Table", "CDM",  "Overfit", "$\\widebar{\\omega}$", "SD($\\omega$)"),  # Format headers
  escape = FALSE  # Allow LaTeX math formatting
) %>%
  kable_styling(latex_options = c("hold_position")) %>%  # Keep table in place
  add_header_above(c(" " = 3, "IMV Metrics" = 3)) %>%  # Group columns
  footnote(
    general_title = "Notes:",
    general = c("IMV(IRT,CDM) based on in-sample data.", 
                "$\\omega$ is the IMV(IRT,CDM) for the out-of-sample fold.")
  ) %>%
  group_rows(index = table(df.imv.estimator$dataname), escape = FALSE)  # Remove bold formatting

# Print LaTeX code
cat(table_latex)
```

### IMV(2PL, CDM)

```{r}
df.imv.compare.combined <- df.imv.compare.combined.raw %>% 
  dplyr :: select(-c(mean_imv_3, sd_imv_3)) %>% 
  pivot_longer(cols = c(mean_imv_1, sd_imv_1, mean_imv_2, sd_imv_2)) %>% 
  filter(!is.na(value)) %>% 
  mutate(estimate_method = ifelse(str_sub(name, -1) == "1", "MAP", "mp")) %>% 
  mutate(name = ifelse(substring(name, 1, 2) == "sd", "sd", "mean")) %>% 
   pivot_wider(names_from = c(name, sample_method), values_from = value ) %>% 
  arrange(dataname, cdm_model, estimate_method) %>% 
  relocate(dataname, cdm_model, estimate_method)

#write_csv(df.imv.compare.combined, "../empirical_output/empirical_output_imv_irt_cdm.csv")
```

```{r height = 8}
df_long <- df.imv.compare.combined %>%
  pivot_longer(cols = c(mean_all, mean_kfold),
               names_to = "method",
               values_to = "mean_imv")

df_long_modified <- df_long %>% 
  mutate(method = ifelse(method == "mean_all", "in-sample", "out-of-sample fold")) %>% 
  rename(Method = method) 


df_error <- df_long_modified %>%
  mutate(sd = ifelse(Method == "out-of-sample fold", df.imv.compare.combined$sd_kfold, 0))
```


```{r}
df_error_update <- df_error %>% 
  mutate(sd = ifelse(Method == "in-sample", NA, sd))

ggplot(df_long_modified, aes(x = dataname, y = mean_imv, fill = Method)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.8)) +
  geom_errorbar(
    data = df_error_update,
    aes(ymin = mean_imv - (sd / sqrt(5)), ymax = mean_imv + (sd / sqrt(5)), x = dataname),
    width = 0.5,
    position = position_dodge(width = 0.8)
  ) + 
  geom_hline(yintercept = 0, linetype = "solid", size = 0.5, color = "black") +  # Emphasized y = 0 line
   facet_grid(estimate_method ~ cdm_model ) +
  labs(x = "Dataset", y = "IMV(2PL, CDM)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.spacing = unit(1.5, "lines"), # Wider space between panels
    axis.title = element_text(s = 15),
    legend.title = element_text(s = 15),
    axis.text = element_text(s = 12),
    strip.text = element_text(s = 15),
    legend.text = element_text(s = 15),
  ) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) + coord_cartesian(ylim = c(-0.05, 0.05))

ggsave("../plots_update/empirical_analysis.pdf", width = 8, height = 6)
```


```{r}
ggplot(df_long_modified %>% filter(dataname == "frac20"), aes(x = dataname, y = mean_imv, fill = Method)) +
  geom_bar(stat = "identity", width = 0.2, position = position_dodge(width = 0.2)) +
  geom_hline(yintercept = 0, linetype = "solid", size = 0.5, color = "black") +  # Emphasized y = 0 line
   facet_grid(estimate_method ~ cdm_model ) +
  geom_errorbar(
    data = df_error_update %>% filter(dataname == "frac20"),
    aes(ymin = mean_imv - sd, ymax = mean_imv + sd, x = dataname),
    width = 0.1,
    position = position_dodge(width = 0.2)
  ) +
 #facet_grid(. ~ estimate_method + cdm_model, scales = "fixed") +
  labs(x = "Dataset", y = "IMV(2PL, CDM)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    
    axis.title = element_text(s = 15),
    legend.title = element_text(s = 15),
    axis.text = element_text(s = 15),
    strip.text = element_text(s = 15),
    legend.text = element_text(s = 15),
    panel.spacing = unit(1.5, "lines")  # Wider space between panels
  ) +
  scale_x_discrete(expand = expansion(mult = c(0.05, 0.05))) + coord_cartesian(ylim = c(-0.05, 0.05))
```

## Check properties of ED
```{r}
obj_names <- c()
n_p <- c()
n_i <- c()
n_p_i <- c()
n_q <- c()
for (obj_name in names(q_data)) {
  df <- q_data[[obj_name]]$to_data_frame()
  # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      if ("booklet" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "booklet"]
      }
  
  n_q <- c(n_q, (length(colnames(df)) - 3))
  obj_names <- c(obj_names, obj_name)
  n_p <- c(n_p,length(unique(df$id)) )
  n_i <- c(n_i,length(unique(df$item)))
  hi <- df %>% 
    group_by(id) %>% 
    tally()
  n_p_i <- c(n_p_i,mean(hi$n))
  
}
```
```{r}
empirical_data_descp <- tibble(datanames = obj_names, n_participants = n_p, n_items = n_i, n_attributes = n_q) %>% 
  arrange(n_attributes) %>% 
  filter(!datanames %in% c("hmcdm_spatialreasoning", "cdm_hr", "cdm_pisa00R"))
```
```{r}
table_latex <- kable(empirical_data_descp, format = "latex", caption = "IRW empirical datasets descriptions") %>%
  kable_styling(latex_options = c("hold_position"))

# Print LaTeX Code
cat(table_latex)
```


## LaTeX table
```{r}
df.imv.compare.combined <- read_csv("../empirical_output/empirical_output_imv_irt_cdm.csv")
```


```{r}
# Generate LaTeX Table with Styling
table_latex <- kable(df.imv.compare.combined, format = "latex", caption = "IMV(2PL, CDM) from IRW empirical datasets") %>%
  kable_styling(latex_options = c("hold_position"))


# Sort data by dataname and round numeric columns to 4 decimals
df_sorted <- df.imv.compare.combined %>%
  arrange(dataname) %>%
  mutate(across(where(is.numeric), round, 5)) %>%
mutate(dataname = ifelse(duplicated(dataname), "", dataname))
  #dplyr :: select(-dataname)  # Remove dataname from table, but keep for grouping

# Create LaTeX table with midrules between dataname groups
table_latex <- kable(
  df_sorted, 
  format = "latex", 
  booktabs = TRUE,  # Ensures \toprule, \midrule, \bottomrule
  caption = "IMV(2PL, CDM) from IRW empirical datasets",
  col.names = c("Table", "CDM", "EstimateMethod", "Overfit", "$\\widebar{\\omega}$", "SD($\\omega$)"),  # Format headers
  escape = FALSE  # Allow LaTeX math formatting
) %>%
  kable_styling(latex_options = c("hold_position")) %>%  # Keep table in place
  add_header_above(c(" " = 3, "IMV Metrics" = 3)) %>%  # Group columns
  footnote(
    general_title = "Notes:",
    general = c("IMV(IRT,CDM) based on in-sample data.", 
                "$\\omega$ is the IMV(IRT,CDM) for the out-of-sample fold.")
  ) %>%
  group_rows(index = table(df.imv.compare.combined$dataname), escape = FALSE)  # Remove bold formatting

# Print LaTeX code
cat(table_latex)
```




