---
title: "IRW explore"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    
---

# Data Preprocessing
## Load Packages
Let's first load the packages that we need for this chapter. 

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
library(dplyr)
library(mirt)
library('ggpubr')
library(Metrics)
library(hash)
library(ggplot2)
library(GGally)
library(gridExtra)
library(rstudioapi)
library(TAM)
library(mgcv)
library(GDINA)
library(MASS)
library(imv)
```
```{r}
#remotes::install_github("hansorlee/irwpkg")
source("bd/00funs.R")
```

```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")

# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```


# Load IRW
```{r}
dataset <- redivis::user("datapages")$dataset("item_response_warehouse",version='v10.0')

dataset_tables <- dataset$list_tables()
```

## Get CDM data
```{r}
##
print(length(dataset_tables))
names<-sapply(dataset_tables,function(x) x$name)
ii<-grep("metadata",names)
names(dataset_tables)<-names
if (length(ii)>0) dataset_tables<-dataset_tables[-ii]   
##
f<-function(table) table$list_variables()
nms<-lapply(dataset_tables,f)

##
f<-function(x) {
    nm<-sapply(x,function(x) x$name)
    any(grepl("^Qmat",nm))
}
test<-sapply(nms,f)
q_data<-dataset_tables[test]
print(length(q_data))
```


# helper functions 
## IRT
```{r}
prob.irt <-function(a, b,theta) {
  1 / (1 + exp(-a * (theta - b)))
}
```

## Convert to Q matrix and data 

```{r}
func.convert.Q.data.structure.with.held.out <- function(df, heldout_rate){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  df.held.out <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = heldout_rate)) %>% 
     dplyr ::arrange(id)
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
     dplyr ::select(-c(held_out)) %>% 
    pivot_wider(names_from = item, values_from = resp) %>% 
     dplyr ::arrange(id) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- df.dat %>%  dplyr ::select(-id)
  tables_list[['heldout']] <- df.held.out
  tables_list[['id']] <- df.dat$id
  
  return (tables_list)
}

func.convert.Q.data.structure.with.k.fold <- function(df, kfold){
  
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  Q <- Q[, colSums(Q != 0) > 0]
  
  Q <- Q[, order(colnames(Q))]
  
  assignments <- sample(rep(1:kfold, length.out = length(df$id)))
  
  df.kfold <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    add_column(gr = assignments) %>% 
     dplyr ::arrange(id)
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat_kfold']] <- df.kfold
  
  return (tables_list)
}

```


## OOS prediction
```{r}
func.run.analysis.kfold <- function(df, cdm_model = "DINA", irt_model = "2PL", if_oos = TRUE, kfold = 5, cdm_pr_estimate = cdm.pr.marg.update, estmethod = "mp"){
  n_items = n_distinct(df$item)
  item_list <- sort(unique(df$item))
  
  df.cdm.structure <- func.convert.Q.data.structure.with.k.fold(df, kfold = kfold)
  
  func.get.prediction.result <- function(df.tmp, dat, qm){
    
    df.ids <- df.tmp %>% group_by(id) %>% 
      tally()
   
    df.dat <- dat %>% 
      dplyr ::select(-gr) %>% 
      dplyr :: arrange(item) %>% 
      pivot_wider(names_from = item, values_from = resp) %>% 
      dplyr ::arrange(id) 
    
    df.dat.no.id <- df.dat %>% dplyr :: select(-id)
    
    # IRT model
    mod.irt <- mirt(df.dat.no.id, 1, itemtype = irt_model)
   
    thetas <- as.data.frame(fscores(mod.irt))

    params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
    
    # CDM model
    predict.matrix <- cdm_pr_estimate(df.dat.no.id, qm, model = cdm_model, estmethod = estmethod)
    
    result <- as.data.frame(predict.matrix) %>% 
      set_names(item_list) %>% 
      mutate(id = df.ids$id) %>% 
     # mutate(id = row_number()) %>% 
      pivot_longer(cols = item_list, names_to = "item", values_to = "prediction.cdm") %>%
      left_join(df.tmp)
    
    result.2 <- tibble(th = thetas$F1, id = df.dat$id) %>%
      left_join(df.tmp) %>% 
      left_join(rownames_to_column(params, var = "item")) %>% 
      mutate(prediction.irt = prob.irt(items.a, items.b, th)) %>% 
      dplyr ::select(id, item, resp, gr, prediction.irt)
    
    result.compared <- result %>% 
      left_join(result.2) 
  }
  
  list.imv <- list()
  
  list.cor <- list()
  
  if (!if_oos) {
    # full data set
    result.compared <- func.get.prediction.result(df.cdm.structure$dat_kfold, df.cdm.structure$dat_kfold, df.cdm.structure$Q)
    
    df.compare <- result.compared %>% 
        filter(!is.na(prediction.irt), !is.na(prediction.cdm))
      
    imv <- imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm)
    
    print(imv)
    
    cor <- cor(df.compare$prediction.irt, df.compare$prediction.cdm)
      
    list.imv <- c(list.imv, imv) 
    
    list.cor <- c(list.cor, cor)
    
  }
  else {
    for (i in seq(1, kfold, 1)) {
      df.tmp <- df.cdm.structure$dat_kfold
      
      df.train <- df.tmp %>% 
        filter(gr != i) 
      
      result.compared <- func.get.prediction.result(df.tmp, df.train, df.cdm.structure$Q)
      
      df.compare <- result.compared %>% 
        filter(gr == i) %>% 
        filter(!is.na(prediction.irt), !is.na(prediction.cdm))
      
      imv <- imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm)
      
      print(imv)
      
      cor <- cor(df.compare$prediction.irt, df.compare$prediction.cdm)
      
      list.imv <- c(list.imv, imv) 
      
      list.cor <- c(list.cor, cor)
    }
  }
  
  return(c(mean(unlist(list.imv)),  sd(unlist(list.imv)), mean(unlist(list.cor)), sd(unlist(list.cor))))
 
}
```


# Empirical analysis



## Organize all data

```{r}
func.run.empirical.data <- function(q_data, cdm_model = "DINA", irt_model = "2PL", cdm_pr_estimate = cdm.pr.marg.update, estmethod = "mp") {
  # Initialize empty lists to store results
  all_results <- list()
  oos_results <- list()
  name_results <- c()
  
  # Iterate over objects in q_data (adjust the loop as needed for your actual structure)
  for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa", "cdm_hr")) {
      print(obj_name)
      
      # Convert the object to a data frame
      #df <- q_data[[obj_name]]$to_data_frame()
      
      if (is.data.frame(q_data[[obj_name]])) {
  df <- q_data[[obj_name]]
} else {
  df <- q_data[[obj_name]]$to_data_frame()  # Assuming the object has this method
}
      
      # Initialize lists to store results for this object
      object_1_results <- list()
      object_2_results <- list()
      
      # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      if ("booklet" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "booklet"]
      }
      
      # Run the first type of analysis once
      object_1_results <- func.run.analysis.kfold(df, cdm_model, irt_model, FALSE, 5, cdm_pr_estimate, estmethod)
      
      # Run the second type of analysis multiple times
      object_2_results <- func.run.analysis.kfold(df, cdm_model, irt_model, TRUE, 5, cdm_pr_estimate, estmethod)
    
      # Combine results for this object and store them in lists
      all_results<- rbind(all_results, object_1_results)
      oos_results <- rbind(oos_results, object_2_results)
      name_results <- c(name_results, obj_name)
      
    }
  }
  
  colnames(all_results) <- c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")
  colnames(oos_results) <- c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")
  
  df.imv.result <- as.data.frame(all_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "all") %>% 
    rbind(as.data.frame(oos_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "kfold")) %>% 
    mutate(mean_imv = as.numeric(mean_imv), 
         sd_imv = as.numeric(sd_imv), 
         mean_cor = as.numeric(mean_cor), 
         sd_cor = as.numeric(sd_cor))
  
  rownames(df.imv.result) <- NULL
    
  # Return the comparison data frame
  return(df.imv.result)
}
```


### irw-data
```{r}
set.seed(10)
```

### non-irw data
timss data 
```{r}
timss2011 <- readRDS("data/TIMSS2011data.rds")

timss_data <-timss2011$data 

item_names <- colnames(timss_data)

timss_data_with_id <- rownames_to_column(timss_data, var = "id")

timss_q <- timss2011$Qc

timss_q_clean <- timss_q[-12, ][,-1][,-1]

timss2011_ready <- timss_data_with_id %>% 
  pivot_longer(cols = -id, names_to = "item", values_to = "resp") %>% 
  left_join(as.data.frame(timss_q_clean) %>% 
  add_column(item = item_names)) %>% 
  filter(resp!= 2)
  
```


mental health data
```{r}
load("data/CDM_AlcoholApplication_Data_IRW.RData")

mental_health_ready <- df %>% 
  rename(resp = "response", V1 = "Qmatrix_skill1", V2 = "Qmatrix_skill2", V3 = "Qmatrix_skill3", V4 = "Qmatrix_skill4")
```

roar pa data
```{r}
load("data/ROAR_PA.RData")

roar_pa_ready <- df %>% 
  dplyr ::select(-rt) %>% 
  mutate(id = as.character(id))
```


```{r}
convert.fraction.data <- function(dat, Q) {
  
  item_names <- colnames(dat)
  
  dat_with_id <- rownames_to_column(dat, var = "id")
  
  dat_ready <- dat_with_id %>%
    pivot_longer(cols = -id, names_to = "item", values_to = "resp") %>%
    left_join(as.data.frame(Q) %>% 
                add_column(item = item_names)) 
  return (dat_ready)
}
```

```{r}
q_data2 <- list(
  timss_2011 = timss2011_ready, 
  mental_health = mental_health_ready,
  roar_pa = roar_pa_ready,
  frac15q5 = convert.fraction.data(data.fraction1$data, data.fraction1$q.matrix),
  frac11q3 = convert.fraction.data(data.fraction2$data, data.fraction2$q.matrix1),
  frac11q5 = convert.fraction.data(data.fraction2$data, data.fraction2$q.matrix2)
)
```



```{r}
empirical_q_data <- append(q_data, q_data2)
```


## sanity check 

### Q1: check if q_matrix is correct?

get from irw
```{r}
df_frac20 <- q_data$frac20$to_data_frame()

df_frac20_clean <- df_frac20 %>% mutate(item = as.integer(substring(item, 6)))

df_frac20_cdm_structure <- func.convert.Q.data.structure.with.k.fold(df_frac20_clean, 0)
```

get from gdina
```{r}
df_frac20_cdm_structure_gdina <- GDINA::frac20
```

```{r}
all(df_frac20_cdm_structure$Q == df_frac20_cdm_structure_gdina$Q)
```


### Q2: check if frac20 cdm prediction makes sense?

```{r}
func.run.analysis.kfold(timss2011_ready, "GDINA", "2PL", FALSE, 5, cdm.pr.marg.update, "mp")
```
```{r}
func.run.analysis.kfold(roar_pa_ready, "GDINA", "2PL", TRUE, 5, cdm.pr.marg.update, "mp")
```

```{r}
df_frac20_analysis <- func.run.analysis.kfold(df_frac20, "GDINA", "2PL", FALSE, 5, cdm.pr.marg.update, "mp")
```
```{r}
df_frac20_analysis <- func.run.analysis.kfold(df_frac20, "GDINA", "2PL", TRUE, 5, cdm.pr.marg.update)
```

### Q3: check identifiability of all data 
```{r}

for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa", "cdm_hr")) {
      print(obj_name)
      df <- q_data[[obj_name]]$to_data_frame() 
      structure <- func.convert.Q.data.structure.with.k.fold(df, 0)
      print(check.qm.omplete(structure$Q))
    }
}

for (obj_name in names(q_data2)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa", "cdm_hr")) {
      print(obj_name)
      df <- q_data2[[obj_name]]
      structure <- func.convert.Q.data.structure.with.k.fold(df, 0)
      print(check.qm.omplete(structure$Q))
    }
}
  
```

```{r}
roar_pa_ready %>% 
  group_by(item) %>% 
  tally()
```

## Run all data
```{r}
set.seed(10)
df.imv.compare.1.mp <- func.run.empirical.data(empirical_q_data, cdm_model = "DINA", irt_model = "2PL")
df.imv.compare.2.mp <- func.run.empirical.data(empirical_q_data, cdm_model = "GDINA", irt_model = "2PL")
df.imv.compare.1.map <- func.run.empirical.data(empirical_q_data, cdm_model = "DINA", irt_model = "2PL", estmethod = "MAP")
df.imv.compare.2.map <- func.run.empirical.data(empirical_q_data, cdm_model = "GDINA", irt_model = "2PL", estmethod = "MAP")
```

```{r}
df.imv.compare.combined <- df.imv.compare.1.mp %>% 
  mutate(cdm_model = "DINA") %>% 
  rbind(df.imv.compare.2.mp %>% mutate(cdm_model = "GDINA")) %>% 
  relocate(dataname, cdm_model) %>% 
  pivot_wider(names_from = "sample_method", values_from = c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")) %>% 
  arrange(dataname) %>% 
  dplyr:: select(dataname, cdm_model, "mean_imv_all", "mean_imv_kfold", "sd_imv_kfold") %>% 
  add_column(estimate_method = "mp") %>% 
  rbind(df.imv.compare.1.map %>% 
  mutate(cdm_model = "DINA") %>% 
  rbind(df.imv.compare.2.map %>% mutate(cdm_model = "GDINA")) %>% 
  relocate(dataname, cdm_model) %>% 
  pivot_wider(names_from = "sample_method", values_from = c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")) %>% 
  arrange(dataname) %>% 
  dplyr:: select(dataname, cdm_model, "mean_imv_all", "mean_imv_kfold", "sd_imv_kfold") %>% 
  add_column(estimate_method = "MAP")) %>% 
  arrange(dataname, cdm_model, estimate_method) %>% 
  relocate(dataname, cdm_model, estimate_method)
```
```{r}
df.imv.compare.combined
```


```{r}
write_csv(df.imv.compare.combined, "analysis/e_analysis_summary.csv")
```

```{r height = 8}
df_long <- df.imv.compare.combined %>%
  pivot_longer(cols = c(mean_imv_all, mean_imv_kfold),
               names_to = "method",
               values_to = "mean_imv")

# Create error bar data (set `sd = 0` for mean_imv_all)
df_error <- df_long %>%
  mutate(sd = ifelse(method == "mean_imv_kfold", df.imv.compare.combined$sd_imv_kfold, 0))

# Plot with all bars having error bars (but `mean_imv_all` has zero error)
ggplot(df_long , aes(x = dataname, y = mean_imv, fill = method)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.8)) +  # Adjust dodge width for alignment
  geom_errorbar(data = df_error, 
                aes(ymin = mean_imv - sd, ymax = mean_imv + sd, x = dataname),
                width = 0.5, 
                position = position_dodge(width = 0.8)) +  # Align error bars with bars
  labs(x = "Dataset", y = "IMV(2PL, CDM)", title = "Comparision of in-sample data vs out-of-sample fold") +
  theme_minimal() +
  facet_grid(estimate_method ~ cdm_model , scales = "free") +
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"  # Move legend to bottom
  ) + scale_x_discrete(expand = expansion(mult = c(0.1, 0.1)))

ggplot(df_long %>% filter(estimate_method == "MAP"), aes(x = dataname, y = mean_imv, fill = method)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.8)) +  # Adjust dodge width for alignment
  geom_errorbar(data = df_error %>% filter(estimate_method == "MAP"), 
                aes(ymin = mean_imv - sd, ymax = mean_imv + sd, x = dataname),
                width = 0.5, 
                position = position_dodge(width = 0.8)) +  # Align error bars with bars
  labs(x = "Dataset", y = "IMV(2PL, CDM)", title = "Comparision of in-sample data vs out-of-sample fold ('MAP' method)") +
  theme_minimal() +
  facet_wrap(~ cdm_model) +
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"  # Move legend to bottom
  ) + scale_x_discrete(expand = expansion(mult = c(0.1, 0.1)))

ggplot(df_long %>% filter(estimate_method == "mp"), aes(x = dataname, y = mean_imv, fill = method)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.8)) +  # Adjust dodge width for alignment
  geom_errorbar(data = df_error %>% filter(estimate_method == "mp"), 
                aes(ymin = mean_imv - sd, ymax = mean_imv + sd, x = dataname),
                width = 0.5, 
                position = position_dodge(width = 0.8)) +  # Align error bars with bars
  labs(x = "Dataset", y = "IMV(2PL, CDM)", title = "Comparision of in-sample data vs out-of-sample fold ('mp' method)") +
  theme_minimal() +
  facet_wrap(~ cdm_model) +
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"  # Move legend to bottom
  ) + scale_x_discrete(expand = expansion(mult = c(0.1, 0.1)))

```
```{r}
df_long_modified <- df_long %>% 
  mutate(method = ifelse(method == "mean_imv_all", "in-sample", "out-of-sample fold")) %>% 
  rename(Method = method) 


df_error <- df_long_modified %>%
  mutate(sd = ifelse(Method == "out-of-sample fold", df.imv.compare.combined$sd_imv_kfold, 0))
```


```{r}
df_error_update <- df_error %>% 
  mutate(sd = ifelse(Method == "in-sample", NA, sd))

ggplot(df_long_modified, aes(x = dataname, y = mean_imv, fill = Method)) +
  geom_bar(stat = "identity", width = 0.8, position = position_dodge(width = 0.8)) +
  geom_errorbar(
    data = df_error_update,
    aes(ymin = mean_imv - sd, ymax = mean_imv + sd, x = dataname),
    width = 0.5,
    position = position_dodge(width = 0.8)
  ) + 
  geom_hline(yintercept = 0, linetype = "solid", size = 0.5, color = "black") +  # Emphasized y = 0 line
   facet_grid(estimate_method ~ cdm_model ) +
 #facet_grid(. ~ estimate_method + cdm_model, scales = "fixed") +
  labs(x = "Dataset", y = "IMV(2PL, CDM)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.spacing = unit(1.5, "lines"), # Wider space between panels
    axis.title = element_text(s = 15),
    legend.title = element_text(s = 10),
    axis.text = element_text(s = 10),
    strip.text = element_text(s = 15),
    legend.text = element_text(s = 10),
  ) +
  scale_x_discrete(expand = expansion(mult = c(0.1, 0.1))) + coord_cartesian(ylim = c(-0.05, 0.05))

ggsave("plots/empirical_analysis_2.pdf", width = 8, height = 6)
```
```{r}
df_error_update <- df_error %>% 
  mutate(sd = ifelse(Method == "in-sample", NA, sd))
```

```{r}
ggplot(df_long_modified %>% filter(dataname == "frac20"), aes(x = dataname, y = mean_imv, fill = Method)) +
  geom_bar(stat = "identity", width = 0.2, position = position_dodge(width = 0.2)) +
  geom_hline(yintercept = 0, linetype = "solid", size = 0.5, color = "black") +  # Emphasized y = 0 line
   facet_grid(estimate_method ~ cdm_model ) +
  geom_errorbar(
    data = df_error_update %>% filter(dataname == "frac20"),
    aes(ymin = mean_imv - sd, ymax = mean_imv + sd, x = dataname),
    width = 0.1,
    position = position_dodge(width = 0.2)
  ) +
 #facet_grid(. ~ estimate_method + cdm_model, scales = "fixed") +
  labs(x = "Dataset", y = "IMV(2PL, CDM)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    
    axis.title = element_text(s = 15),
    legend.title = element_text(s = 15),
    axis.text = element_text(s = 15),
    strip.text = element_text(s = 15),
    legend.text = element_text(s = 15),
    panel.spacing = unit(1.5, "lines")  # Wider space between panels
  ) +
  scale_x_discrete(expand = expansion(mult = c(0.05, 0.05))) + coord_cartesian(ylim = c(-0.05, 0.05))
```

```{r}
# ggplot(df_long %>% filter(dataname == "frac20"), aes(x = dataname, y = mean_imv, fill = method)) +
#   geom_bar(stat = "identity", width = 0.5, position = position_dodge(width = 0.5)) +  # Adjust dodge width for alignment
#   geom_errorbar(data = df_error %>% filter(dataname == "frac20"), 
#                 aes(ymin = mean_imv - sd, ymax = mean_imv + sd, x = dataname),
#                 width = 0.25, 
#                 position = position_dodge(width = 0.5)) +  # Align error bars with bars
#   labs(x = "Dataset", y = "IMV(2PL, CDM)", title = "Comparision of in-sample data vs out-of-sample fold") +
#   theme_minimal() +
#   facet_grid(estimate_method ~ cdm_model) +
#    theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     legend.position = "bottom"  # Move legend to bottom
#   ) + ylim(-0.2, 0.05)
# 
# ggplot(df_long %>% filter(dataname == "frac20", method != "mean_imv_kfold"), aes(x = dataname, y = mean_imv, fill = method)) +
#   geom_bar(stat = "identity", width = 0.25, position = position_dodge(width = 0.25)) +  # Adjust dodge width for alignment
#   
#   labs(x = "Dataset", y = "IMV(2PL, CDM)", title = "in-sample data prediction") +
#   theme_minimal() +
#   facet_grid(estimate_method ~ cdm_model) +
#    theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     legend.position = "bottom"  # Move legend to bottom
#   ) + ylim(-0.05, 0.05)
```

## Check properties of ED
```{r}
obj_names <- c()
n_p <- c()
n_i <- c()
n_p_i <- c()
n_q <- c()
for (obj_name in names(q_data)) {
  df <- q_data[[obj_name]]$to_data_frame()
  # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      if ("booklet" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "booklet"]
      }
  
  n_q <- c(n_q, (length(colnames(df)) - 3))
  obj_names <- c(obj_names, obj_name)
  n_p <- c(n_p,length(unique(df$id)) )
  n_i <- c(n_i,length(unique(df$item)))
  hi <- df %>% 
    group_by(id) %>% 
    tally()
  n_p_i <- c(n_p_i,mean(hi$n))
  
}
```
```{r}
empirical_data_descp <- tibble(datanames = obj_names, n_participants = n_p, n_items = n_i, n_attributes = n_q) %>% 
  arrange(n_attributes) %>% 
  filter(!datanames %in% c("hmcdm_spatialreasoning", "cdm_hr", "cdm_pisa00R"))
```
```{r}
table_latex <- kable(empirical_data_descp, format = "latex", caption = "IRW empirical datasets descriptions") %>%
  kable_styling(latex_options = c("hold_position"))

# Print LaTeX Code
cat(table_latex)
```

### check the structure of Qmatrix
```{r}
obj_name = "cdm_ecpe"
df <- q_data[[obj_name]]$to_data_frame() 
```

```{r}
q_matrix <- func.convert.Q.data.structure.with.k.fold(df, 5)$Q
```

```{r}
mean(rowMeans(frac20$Q))
sd(rowMeans(frac20$Q))
```

```{r}
q_matrix <- as.matrix(q_matrix)  # Convert to matrix
q_matrix <- apply(q_matrix, 2, as.numeric)
co_occurrence <- t(q_matrix) %*% q_matrix
#print(co_occurrence)

# Convert adjacency matrix into an edge list
edges <- which(adj_matrix > 0, arr.ind = TRUE)
print(edges)
# Create a data frame of edges (directed from colnames[edges[,1]] → colnames[edges[,2]])
edge_list <- data.frame(
  from = colnames(q_matrix)[edges[, 1]],
  to = colnames(q_matrix)[edges[, 2]],
  weight = adj_matrix[edges]  # Edge weights based on co-occurrence
) %>% filter(!is.na(from), !is.na(to))

print(edge_list)  # View edges

# Create directed graph using igraph
g <- graph_from_data_frame(edge_list, directed = TRUE)

# Plot the directed graph using ggraph
ggraph(g, layout = "fr") +  # "fr" = Force-directed layout
  geom_edge_link(aes(edge_alpha = weight), arrow = arrow(length = unit(4, 'mm')), end_cap = circle(3, 'mm')) +
  geom_node_point(size = 4, color = "red") +  # Node style
  geom_node_text(aes(label = name), vjust = 1.5, size = 4) +  # Label nodes
  theme_void()  # Clean theme
```

## LaTeX table
```{r}
library(kableExtra)
```

```{r}
# Generate LaTeX Table with Styling
table_latex <- kable(df.imv.compare.combined, format = "latex", caption = "IMV(2PL, CDM) from IRW empirical datasets") %>%
  kable_styling(latex_options = c("hold_position"))


# Sort data by dataname and round numeric columns to 4 decimals
df_sorted <- df.imv.compare.combined %>%
  arrange(dataname) %>%
  mutate(across(where(is.numeric), round, 5)) %>%
mutate(dataname = ifelse(duplicated(dataname), "", dataname))
  #dplyr :: select(-dataname)  # Remove dataname from table, but keep for grouping

# Create LaTeX table with midrules between dataname groups
table_latex <- kable(
  df_sorted, 
  format = "latex", 
  booktabs = TRUE,  # Ensures \toprule, \midrule, \bottomrule
  caption = "IMV(2PL, CDM) from IRW empirical datasets",
  col.names = c("Table", "CDM", "EstimateMethod", "Overfit", "$\\widebar{\\omega}$", "SD($\\omega$)"),  # Format headers
  escape = FALSE  # Allow LaTeX math formatting
) %>%
  kable_styling(latex_options = c("hold_position")) %>%  # Keep table in place
  add_header_above(c(" " = 3, "IMV Metrics" = 3)) %>%  # Group columns
  footnote(
    general_title = "Notes:",
    general = c("IMV(IRT,CDM) based on in-sample data.", 
                "$\\omega$ is the IMV(IRT,CDM) for the out-of-sample fold.")
  ) %>%
  group_rows(index = table(df.imv.compare.combined$dataname), escape = FALSE)  # Remove bold formatting

# Print LaTeX code
cat(table_latex)
```




