---
title: "IRW explore"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    
---

# Data Preprocessing
## Load Packages
Let's first load the packages that we need for this chapter. 

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
library(dplyr)
library(mirt)
library('ggpubr')
library(Metrics)
library(hash)
library(ggplot2)
library(GGally)
library(gridExtra)
library(rstudioapi)
library(TAM)
library(mgcv)
library(GDINA)
library(MASS)
library(imv)
```
```{r}
source("bd/00funs.R")
```

```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")

# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

# Load IRW
```{r}
dataset <- redivis::user("datapages")$dataset("item_response_warehouse",version='v10.0')

dataset_tables <- dataset$list_tables()
```

## Get CDM data
```{r}
##
print(length(dataset_tables))
names<-sapply(dataset_tables,function(x) x$name)
ii<-grep("metadata",names)
names(dataset_tables)<-names
if (length(ii)>0) dataset_tables<-dataset_tables[-ii]   
##
f<-function(table) table$list_variables()
nms<-lapply(dataset_tables,f)

##
f<-function(x) {
    nm<-sapply(x,function(x) x$name)
    any(grepl("^Qmat",nm))
}
test<-sapply(nms,f)
q_data<-dataset_tables[test]
print(length(q_data))
```


# helper functions 
## IRT
```{r}
prob.irt <-function(a, b,theta) {
  1 / (1 + exp(-a * (theta - b)))
}
```

## Convert to Q matrix and data 

```{r}
func.convert.Q.data.structure.with.held.out <- function(df, heldout_rate){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  df.held.out <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = heldout_rate)) %>% 
     dplyr ::arrange(id)
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
     dplyr ::select(-c(held_out)) %>% 
    pivot_wider(names_from = item, values_from = resp) %>% 
     dplyr ::arrange(id) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- df.dat %>%  dplyr ::select(-id)
  tables_list[['heldout']] <- df.held.out
  tables_list[['id']] <- df.dat$id
  
  return (tables_list)
}
```

```{r}
func.convert.Q.data.structure.with.k.fold <- function(df, kfold){
  
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  Q <- Q[, colSums(Q != 0) > 0]
  
  Q <- Q[, order(colnames(Q))]
  
  assignments <- sample(rep(1:kfold, length.out = length(df$id)))
  
  df.kfold <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    add_column(gr = assignments) %>% 
     dplyr ::arrange(id)
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat_kfold']] <- df.kfold
  
  return (tables_list)
}

```


## OOS prediction
```{r}
func.run.analysis.kfold <- function(df, cdm_model = "DINA", irt_model = "2PL", if_oos = TRUE, kfold = 5, cdm_pr_estimate = cdm.pr.marg){
  n_items = n_distinct(df$item)
  item_list <- sort(unique(df$item))
  
  df.cdm.structure <- func.convert.Q.data.structure.with.k.fold(df, kfold = kfold)
  
  func.get.prediction.result <- function(df.tmp, dat, qm){
    
    df.ids <- df.tmp %>% group_by(id) %>% 
      tally()
   
    df.dat <- dat %>% 
      dplyr ::select(-gr) %>% 
      dplyr :: arrange(item) %>% 
      pivot_wider(names_from = item, values_from = resp) %>% 
      dplyr ::arrange(id) 
    
    df.dat.no.id <- df.dat %>% dplyr :: select(-id)
    
    # IRT model
    mod.irt <- mirt(df.dat.no.id, 1, itemtype = irt_model)
   
    thetas <- as.data.frame(fscores(mod.irt))

    params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
    
    # CDM model
    predict.matrix <- cdm_pr_estimate(df.dat.no.id, qm, model = cdm_model)
    
    result <- as.data.frame(predict.matrix) %>% 
      set_names(item_list) %>% 
      mutate(id = df.ids$id) %>% 
     # mutate(id = row_number()) %>% 
      pivot_longer(cols = item_list, names_to = "item", values_to = "prediction.cdm") %>%
      left_join(df.tmp)
    
    result.2 <- tibble(th = thetas$F1, id = df.dat$id) %>%
      left_join(df.tmp) %>% 
      left_join(rownames_to_column(params, var = "item")) %>% 
      mutate(prediction.irt = prob.irt(items.a, items.b, th)) %>% 
      dplyr ::select(id, item, resp, gr, prediction.irt)
    
    result.compared <- result %>% 
      left_join(result.2) 
  }
  
  list.imv <- list()
  
  list.cor <- list()
  
  if (!if_oos) {
    # full data set
    result.compared <- func.get.prediction.result(df.cdm.structure$dat_kfold, df.cdm.structure$dat_kfold, df.cdm.structure$Q)
    
    df.compare <- result.compared %>% 
        filter(!is.na(prediction.irt), !is.na(prediction.cdm))
      
    imv <- imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm)
    
    print(imv)
    
    cor <- cor(df.compare$prediction.irt, df.compare$prediction.cdm)
      
    list.imv <- c(list.imv, imv) 
    
    list.cor <- c(list.cor, cor)
    
  }
  else {
    for (i in seq(1, kfold, 1)) {
      df.tmp <- df.cdm.structure$dat_kfold
      
      df.train <- df.tmp %>% 
        filter(gr != i) 
      
      result.compared <- func.get.prediction.result(df.tmp, df.train, df.cdm.structure$Q)
      
      df.compare <- result.compared %>% 
        filter(gr == i) %>% 
        filter(!is.na(prediction.irt), !is.na(prediction.cdm))
      
      imv <- imv::imv.binary(df.compare$resp, df.compare$prediction.irt, df.compare$prediction.cdm)
      
      print(imv)
      
      cor <- cor(df.compare$prediction.irt, df.compare$prediction.cdm)
      
      list.imv <- c(list.imv, imv) 
      
      list.cor <- c(list.cor, cor)
    }
  }
  
  return(c(mean(unlist(list.imv)),  sd(unlist(list.imv)), mean(unlist(list.cor)), sd(unlist(list.cor))))
 
}
```


# Empirical analysis

## sanity check 

### Q1: check if q_matrix is correct?

get from irw
```{r}
df_frac20 <- q_data$frac20$to_data_frame()

df_frac20_clean <- df_frac20 %>% mutate(item = as.integer(substring(item, 6)))

df_frac20_cdm_structure <- func.convert.Q.data.structure.with.k.fold(df_frac20_clean, 0)
```

get from gdina
```{r}
df_frac20_cdm_structure_gdina <- GDINA::frac20
```

```{r}
all(df_frac20_cdm_structure$Q == df_frac20_cdm_structure_gdina$Q)
```


### Q2: check if frac20 cdm prediction makes sense?

```{r}
df_frac20_analysis <- func.run.analysis.kfold(df_frac20, "GDINA", "2PL", FALSE, 5, cdm.pr.marg)
```
```{r}
df_frac20_analysis <- func.run.analysis.kfold(df_frac20, "GDINA", "2PL", TRUE, 5, cdm.pr.marg)
```
```{r}
df_frac20_analysis
```

## Run all data

```{r}
func.run.empirical.data <- function(cdm_model = "DINA", irt_model = "2PL", cdm_pr_estimate = cdm.pr) {
  # Initialize empty lists to store results
  all_results <- list()
  oos_results <- list()
  name_results <- c()
  
  # Iterate over objects in q_data (adjust the loop as needed for your actual structure)
  for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa")) {
      print(obj_name)
      
      # Convert the object to a data frame
      df <- q_data[[obj_name]]$to_data_frame()
      
      # Initialize lists to store results for this object
      object_1_results <- list()
      object_2_results <- list()
      
      # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      if ("booklet" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "booklet"]
      }
      
      # Run the first type of analysis once
      object_1_results <- func.run.analysis.kfold(df, cdm_model, irt_model, FALSE, 5, cdm.pr.marg)
      
      # Run the second type of analysis multiple times
      object_2_results <- func.run.analysis.kfold(df, cdm_model, irt_model, TRUE, 5, cdm.pr.marg)
    
      # Combine results for this object and store them in lists
      all_results<- rbind(all_results, object_1_results)
      oos_results <- rbind(oos_results, object_2_results)
      name_results <- c(name_results, obj_name)
      
    }
  }
  
  colnames(all_results) <- c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")
  colnames(oos_results) <- c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")
  
  df.imv.result <- as.data.frame(all_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "all") %>% 
    rbind(as.data.frame(oos_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "kfold")) %>% 
    mutate(mean_imv = as.numeric(mean_imv), 
         sd_imv = as.numeric(sd_imv), 
         mean_cor = as.numeric(mean_cor), 
         sd_cor = as.numeric(sd_cor))
  
  rownames(df.imv.result) <- NULL
    
  # Return the comparison data frame
  return(df.imv.result)
}
```


```{r}
func.run.empirical.data.timss <- function(cdm_model = "DINA", irt_model = "2PL", cdm_pr_estimate = cdm.pr) {
  # Initialize empty lists to store results
  all_results <- list()
  oos_results <- list()
  name_results <- c()
  
  # Iterate over objects in q_data (adjust the loop as needed for your actual structure)
  for (obj_name in names(q_data)) {
    if (substring(obj_name, 1, 11) == "cdm_timss07") {
      print(obj_name)
      
      # Convert the object to a data frame
      df <- q_data[[obj_name]]$to_data_frame()
      
      # Initialize lists to store results for this object
      object_1_results <- list()
      object_2_results <- list()
      
      # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      if ("booklet" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "booklet"]
      }
      
      # Run the first type of analysis once
      object_1_results <- func.run.analysis.kfold(df, cdm_model, irt_model, FALSE, 5, cdm.pr.marg)
      
      # Run the second type of analysis multiple times
      object_2_results <- func.run.analysis.kfold(df, cdm_model, irt_model, TRUE, 5, cdm.pr.marg)
    
      # Combine results for this object and store them in lists
      all_results<- rbind(all_results, object_1_results)
      oos_results <- rbind(oos_results, object_2_results)
      name_results <- c(name_results, obj_name)
      
    }
  }
  
  colnames(all_results) <- c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")
  colnames(oos_results) <- c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")
  
  df.imv.result <- as.data.frame(all_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "all") %>% 
    rbind(as.data.frame(oos_results) %>% 
    add_column(dataname = name_results) %>% 
    mutate(sample_method = "kfold")) %>% 
    mutate(mean_imv = as.numeric(mean_imv), 
         sd_imv = as.numeric(sd_imv), 
         mean_cor = as.numeric(mean_cor), 
         sd_cor = as.numeric(sd_cor))
  
  rownames(df.imv.result) <- NULL
    
  # Return the comparison data frame
  return(df.imv.result)
}
```


### 2PL Non-timss
```{r}
set.seed(10)

df.imv.compare.1 <- func.run.empirical.data(cdm_model = "DINA", irt_model = "2PL")

df.imv.compare.2 <- func.run.empirical.data(cdm_model = "DINO", irt_model = "2PL")

df.imv.compare.3 <- func.run.empirical.data(cdm_model = "GDINA", irt_model = "2PL")

```
```{r}
df.imv.compare.result <- df.imv.compare.1 %>% 
  mutate(cdm_model = "DINA") %>% 
  rbind(df.imv.compare.2 %>% mutate(cdm_model = "DINO")) %>% 
  rbind(df.imv.compare.3 %>% mutate(cdm_model = "GDINA")) %>% 
  relocate(dataname, cdm_model) %>% 
  pivot_wider(names_from = "sample_method", values_from = c("mean_imv",  "sd_imv", "mean_cor", "sd_cor")) %>% 
  arrange(dataname) %>% 
  dplyr:: select(dataname, cdm_model, "mean_imv_all", "mean_imv_kfold", "sd_imv_kfold")
 
```

```{r}
df.imv.compare.result
```

### 2PL timss
```{r}
timss2011 <- readRDS("data/TIMSS2011data.rds")
```
```{r}
timss_data <-timss2011$data 

item_names <- colnames(timss_data)

timss_data_with_id <- rownames_to_column(timss_data, var = "id")

  
timss_q <- timss2011$Qc

timss_q_clean <- timss_q[-12, ][,-1][,-1]

timss2011_ready <- timss_data_with_id %>% 
  pivot_longer(cols = -id, names_to = "item", values_to = "resp") %>% 
  left_join(as.data.frame(timss_q_clean) %>% 
  add_column(item = item_names)) %>% 
  filter(resp!= 2)
  
```

```{r}
line <- c()
for (model in c("DINA", "DINO", "GDINA")){
  line <- c(line, func.run.analysis.kfold(timss2011_ready, model, "2PL", FALSE, 5, cdm.pr.marg))
  line <- c(line, func.run.analysis.kfold(timss2011_ready, model, "2PL", TRUE, 5, cdm.pr.marg))
  
}

```
```{r}
line
```
### frac data more things
```{r}
run.fraction.data.analysis <- function(dat, Q) {
  
  item_names <- colnames(dat)
  
  dat_with_id <- rownames_to_column(dat, var = "id")
  
  dat_ready <- dat_with_id %>%
    pivot_longer(cols = -id, names_to = "item", values_to = "resp") %>%
    left_join(as.data.frame(Q) %>% 
                add_column(item = item_names)) 
  hi <- NULL
  for (model in c("DINA", "DINO", "GDINA")){
    hi <- hi %>% rbind(func.run.analysis.kfold(dat_ready, model, "2PL", FALSE, 5, cdm.pr.marg) %>% rbind(func.run.analysis.kfold(dat_ready, model, "2PL", TRUE, 5, cdm.pr.marg)))
    
  }
  return (hi)
}

```
```{r}
run.fraction.data.analysis(data.fraction1$data, data.fraction1$q.matrix)
```
```{r}
run.fraction.data.analysis(data.fraction2$data, data.fraction2$q.matrix1)
```
```{r}
run.fraction.data.analysis(data.fraction2$data, data.fraction2$q.matrix2)
```
```{r}
run.fraction.data.analysis(data.fraction2$data, data.fraction2$q.matrix3)
```
```{r}
run.fraction.data.analysis(data.fraction4$data, data.fraction4$q.matrix %>% dplyr :: select(-item))
```
```{r}
data.fraction5
```

## Check properties of ED
```{r}
obj_names <- c()
n_p <- c()
n_i <- c()
n_p_i <- c()
n_q <- c()
for (obj_name in names(q_data)) {
  df <- q_data[[obj_name]]$to_data_frame()
  
  # Remove "rt" column if it exists
      if ("rt" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "rt"]
      }
      
      if ("booklet" %in% colnames(df)) {
        df <- df[, !colnames(df) %in% "booklet"]
      }
  
  n_q <- c(n_q, (length(colnames(df)) - 3))
  obj_names <- c(obj_names, obj_name)
  n_p <- c(n_p,length(unique(df$id)) )
  n_i <- c(n_i,length(unique(df$item)))
  hi <- df %>% 
    group_by(id) %>% 
    tally()
  n_p_i <- c(n_p_i,mean(hi$n))
  
}
```
```{r}
empirical_data_descp <- tibble(datanames = obj_names, n_participants = n_p, n_items = n_i, n_attributes = n_q) %>% 
  arrange(n_attributes) %>% 
  filter(!datanames %in% c("hmcdm_spatialreasoning", "cdm_pisa00R"))
```
```{r}
table_latex <- kable(empirical_data_descp, format = "latex", caption = "IRW empirical datasets descriptions") %>%
  kable_styling(latex_options = c("hold_position"))

# Print LaTeX Code
cat(table_latex)
```


## 1PL 
```{r}
set.seed(10)

df.imv.compare.1.1pl <- func.run.empirical.data(cdm_model = "DINA", irt_model = "Rasch")

df.imv.compare.2.1pl <- func.run.empirical.data(cdm_model = "DINO", irt_model = "Rasch")

df.imv.compare.3.1pl <- func.run.empirical.data(cdm_model = "GDINA", irt_model = "Rasch")
```

## LaTeX table
```{r}
library(kableExtra)
```

```{r}
# Generate LaTeX Table with Styling
table_latex <- kable(df.imv.compare.result, format = "latex", caption = "IMV(2PL, CDM) from IRW empirical datasets") %>%
  kable_styling(latex_options = c("hold_position"))

# Print LaTeX Code
cat(table_latex)
```
## Parameter Analysis
```{r}
func.convert.Q.data.structure.with.held.out <- function(df, heldout_rate){
  Q <- df %>% 
     dplyr ::select(-c(id, resp)) %>% 
    unique() %>% 
     dplyr ::arrange(item) %>% 
     dplyr ::select(-c(item))
  
  df.held.out <- df %>% 
     dplyr ::select(item, id, resp) %>% 
    dplyr :: arrange(item) %>% 
    mutate(held_out = rbinom(n = n(), size = 1, prob = heldout_rate)) %>% 
     dplyr ::arrange(id)
  
  df.dat <- df.held.out %>% 
    filter(held_out == 0) %>% 
     dplyr ::select(-c(held_out)) %>% 
     dplyr :: arrange(item) %>% 
    pivot_wider(names_from = item, values_from = resp) %>% 
     dplyr ::arrange(id) 
  
  tables_list <- list()
  tables_list[['Q']] <- Q
  tables_list[['dat']] <- df.dat %>%  dplyr ::select(-id)
  tables_list[['heldout']] <- df.held.out
  tables_list[['id']] <- df.dat$id
  
  
  return (tables_list)
}
```

```{r}
df.tmp <- q_data$frac20

df.tmp <- df.tmp$to_data_frame()


```
```{r}
df.cdm.structure <- func.convert.Q.data.structure.with.held.out(df.tmp, 0)
```


```{r}
item_list <- sort(unique(df.tmp$item))

fit1 <- GDINA(dat = df.cdm.structure$dat, Q = df.cdm.structure$Q, model = "DINA")

cdm.params <- as.data.frame(coef(object = fit1, what = "gs")) %>% 
  add_column(item = item_list)

mod.irt <- mirt(df.cdm.structure$dat, 1, itemtype = "2PL")
  
params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
  
irt.params <- params %>% 
    left_join(rownames_to_column(params, var = "item")) 

result <- cdm.params %>% 
  left_join(irt.params)


```

```{r}
get.item.params <- function(q_data){
  result.final <- tibble()
  for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa")) {
      print(obj_name)
      
      # Convert the object to a data frame
      df.tmp <- q_data[[obj_name]]$to_data_frame()
      
      df.cdm.structure <- func.convert.Q.data.structure.with.held.out(df.tmp, 0)
      
      item_list <- sort(unique(df.tmp$item))
      
      fit1 <- GDINA(dat = df.cdm.structure$dat, Q = df.cdm.structure$Q, model = "DINA")
      
      cdm.params <- as.data.frame(coef(object = fit1, what = "gs")) %>% 
        add_column(item = item_list)
      
      mod.irt <- mirt(df.cdm.structure$dat, 1, itemtype = "2PL")
      params <- data.frame(coef(mod.irt, IRTpars = TRUE, simplify = TRUE))
      
      irt.params <- params %>% 
        left_join(rownames_to_column(params, var = "item")) 
      
      result <- cdm.params %>% 
        left_join(irt.params) %>% 
        add_column(dataname = obj_name)
      
      result.final <- result.final %>% rbind(result)
    }
  }
  return (result.final)

}
```


```{r}
df.params.empirical <- get.item.params(q_data)
```

```{r}
df.params.empirical.clean <- df.params.empirical %>% 
  dplyr :: rename(g = guessing, s = slip, a = items.a, b = items.b) %>% 
  dplyr :: select(dataname, item, a, b, g, s)
```


```{r}
create_ggpairs <- function(data, title) {
  ggpairs(
    data,
    columns = c("a", "b", "g", "s"),       # Columns for scatterplot matrix
    mapping = aes(label = item),          # Map 'item' as point label
    title = title
  ) + theme_bw()
}

for (i in unique(df.params.empirical.clean$dataname)) {
  create_ggpairs(df.params.empirical.clean %>% filter(dataname == i), i)
}

```
```{r}
# Store the plots in a list
plot_list <- list()

# Loop through unique datanames and create ggpairs plots
for (i in unique(df.params.empirical.clean$dataname)) {
  plot_list[[i]] <- create_ggpairs(
    df.params.empirical.clean %>% filter(dataname == i),
    i
  )
}


```

```{r}
pdf("plots/empirical_scatterplot_frac20.pdf", width = 5, height = 5)

# Print the plot to the PDF device
print(plot_list$frac20)

dev.off()

pdf("plots/empirical_scatterplot_cdm_hr.pdf", width = 5, height = 5)

# Print the plot to the PDF device
print(plot_list$cdm_hr)

dev.off()

pdf("plots/empirical_scatterplot_mcmi_mokken.pdf", width = 5, height = 5)

# Print the plot to the PDF device
print(plot_list$mcmi_mokken)

dev.off()

pdf("plots/empirical_scatterplot_cdm_ecpe.pdf", width = 5, height = 5)

# Print the plot to the PDF device
print(plot_list$cdm_ecpe)

# Close the PDF device
dev.off()
```
```{r}
for (obj_name in names(q_data)) {
    if (!substring(obj_name, 1, 9) %in% c("cdm_timss", "cdm_pisa0", "hmcdm_spa")) {
      print(obj_name)
      
      # Convert the object to a data frame
      df.tmp <- q_data[[obj_name]]$to_data_frame()
      
      df.cdm.structure <- func.convert.Q.data.structure.with.held.out(df.tmp, 0)
      
      print(length(df.cdm.structure$Q))
      print(length(df.cdm.structure$dat))
    }
  }
```

```{r}

```



